### 2023-08-15 paper_codes.Rmd
Add files via upload

---
title: "Paper Codes"
author: "Markus Viljanen"
date: "2023-08-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
library(plyr)
library(tidyverse)
library(sf)
library(INLA)
library(mgcv)
library(ggplot2)
library(cowplot)
library(colorspace)
library(scales)
theme_set(theme_bw())
```

## Data set

Species to analyze and their Latin names

```{r}
species.map <- read.csv(file="data_paper/chosen_species.csv") %>% transmute(species=Nednaam, name=Wetnaam)
species <- species.map$species
```

Netherlands boundary shapefile

```{r}
NL <- st_read("data_paper/NL.shp")$geometry#st_simplify()
NL <- (NL - matrix(data=c(155000, 463000), ncol=2)) * matrix(data=c(1/1000, 1/1000), ncol=2)
```

Data for vegetation plots and abiotic measurements data

```{r}
# Load field visit data
df.plots <- read.csv(file="data_paper/data_plots.csv", check.names = F)
# Add geometries
sf.plots <- df.plots %>% st_as_sf(coords = c("X","Y"), crs=st_crs(NL))
sf.plots[c("X", "Y")] <- sf.plots  %>% st_coordinates()
```

Data for Netherlands grid

```{r}
# Load Netherlands grid data
df.grid <- read.csv(file="data_paper/data_grid.csv", check.names = F)
# Grid with rectangles and their centers
grid.rect <- st_make_grid(NL,n=100,crs=st_crs(NL), what='polygons', square=T) 
grid.circle <- st_make_grid(NL,n=100,crs=st_crs(NL), what='centers')
index <- which(lengths(st_within(grid.circle, NL)) > 0) #st_intersects(grid.rect, NL)
# Add geometries like originally created
geoms <- data.frame(grid.circle[index], grid.rect[index]) 
sf.grid <- cbind(df.grid, geoms) %>% st_as_sf(crs=st_crs(NL))
```

## Plot field visits

Translate landuse to English

```{r}
sf.grid$landuse <- recode(sf.grid$bodemgebruik, "Bos"="Forest", "Bebouwd"="Built", 
                          "Droog natuurlijk terrein"="Natural terrain (dry)",
                          "Hoofdweg"="Highway","Landbouw en overig agrarisch"="Agricultural", 
                          "Nat natuurlijk terrein"= "Natural terrain (wet)", 
                          "Recreatie"="Recreational", "Spoorterrein"="Railway", "Water"="Water")
sf.grid$landuse <- fct_relevel(sf.grid$landuse, "Forest", "Natural terrain (dry)", "Natural terrain (wet)", 
                               "Agricultural", "Recreational", "Highway", "Railway", "Water", "Built")
sf.plots$landuse <- recode(sf.plots$bodemgebruik, "Bos"="Forest", "Bebouwd"="Built", 
                          "Droog natuurlijk terrein"="Natural terrain (dry)",
                          "Hoofdweg"="Highway","Landbouw en overig agrarisch"="Agricultural", 
                          "Nat natuurlijk terrein"= "Natural terrain (wet)", 
                          "Recreatie"="Recreational", "Spoorterrein"="Railway", "Water"="Water")
sf.plots$landuse <- fct_relevel(sf.plots$landuse, "Forest", "Natural terrain (dry)", "Natural terrain (wet)", 
                               "Agricultural", "Recreational", "Highway", "Railway", "Water", "Built")
```

Plot landuse in each grid cell:

```{r}
ggplot(NL) + geom_sf() +
	geom_sf(aes(geometry=geometry.1, fill=landuse), data=sf.grid) +
  scale_fill_manual(values=c("Forest"="forestgreen", "Built"="slategrey",
                             "Natural terrain (dry)"="peru","Natural terrain (wet)"="lightseagreen",
                             "Highway"="grey","Railway"="seashell2", "Agricultural"="wheat",#
                             "Recreatie"="lawngreen", "Water"="dodgerblue"
                             )) +
	theme(legend.position="right") #+ggtitle('Grid of Netherlands: Landuse') 
ggsave('results_paper/landuse.png', width=1600, heigh=1200, units='px')
```

Bar plot of landuse in grid cells vs. vegetation plots:

```{r}
f.grid <- table(sf.grid$landuse)
f.plots <- table(sf.plots$landuse)
f <- rbind(data.frame(f.grid / sum(f.grid), Data='Grid'),
           data.frame(f.plots / sum(f.plots), Data='Plots'))

ggplot(data=f, aes(x=Var1, y=Freq, fill=Data)) +
  geom_bar(stat="identity", position=position_dodge()) +
  scale_fill_manual(values=c('#999999','#E69F00')) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  xlab('Landuse') + ylab('% of total') #+ ggtitle('Landuse: Grid of Netherlands vs. Field visits')
ggsave('results_paper/landuse_vs_visit.png', width=1400, heigh=900, units='px')
```

Plot field visits to vegetation plots and abiotic measurements over the grid:

```{r}
sf.plots$is_fieldvisit <- !is.na(sf.plots$Dagnummer)
sf.plots$has_abiotics <- !is.na(sf.plots$releve_nr)
```

```{r}
ggplot(NL) + geom_sf() +
	geom_sf(aes(geometry=geometry.1), data=sf.grid) + 
	#geom_sf(aes(geometry=geometry), shape=3, size=0.3, data=sf.grid) + 
	geom_sf(aes(geometry=geometry, color=has_abiotics), shape=4, size=0.5, data=sf.plots) + 
	theme(plot.title = element_text(hjust = 0.5), legend.title=element_text(size=10), 
		legend.text=element_text(size=8), legend.position="right") +
	scale_colour_manual(values=c("#377EB8", "#4DAF4A"), name="Has abiotics") #+ 
	#ggtitle('Location of Field Visits and Abiotics') 
ggsave('results_paper/visits.png', width=1400, heigh=1200, units='px')

```

## Spatial interpolation of common heather

Create a mesh over the Netherlands for spatial autocorrelation modelling

```{r}
step <- 10000 / 1000
domain <- inla.nonconvex.hull(as.matrix(sf.grid[c("X","Y")])[,1:2], convex=-0.04)
mesh <- inla.mesh.2d(boundary=domain, max.edge=c(step, step*4), cutoff=step/4)
```

Illustrate the mesh and common heather (Struikhei) occurrence

```{r}

plot(NL, col="grey", main='Gaussian Markov Random Field')
plot(mesh, asp=1, add=T)
rbPal <- colorRampPalette(c('lightyellow','green'))
col <- rbPal(2)[sf.plots[["Struikhei"]] + 1]
plot(sf.plots["Struikhei"], pch=21, bg=col, pal=rbPal, cex=0.2, add=T, alpha=0.2)
legend(-200,150, text.font=4, cex=0.6, legend=c("Not Observed", "Observed"), col=c("lightyellow", "green"), pch=21)#16000, 614000
```

```{r}
## Prior: log(range) ~N(0, 10), log(sigma) = N(0,10)
#range0 <- 1 #exp(10) # ~22 km
#sigma0 <- 1 # ~2.7
#kappa0 <- sqrt(8)/range0
#tau0 <- 1/(sqrt(4*pi)*kappa0*sigma0) #should be 8?
#spde <- inla.spde2.matern(mesh,alpha=2,
#                          B.tau=matrix(c(log(tau0), -1, +1), nrow=1, ncol=3),
#                          B.kappa=matrix(c(log(kappa0), 0, -1), nrow=1, ncol=3),
#                          theta.prior.mean=c(0,0), theta.prior.prec=c(0.1,0.1))
#s.index <- inla.spde.make.index(name="spatial.field", n.spde=spde$n.spde) 
```

Interpolate common heather occurrence in the Netherlands with simple kriging, save zscore of log-odds as pH indicator:

```{r}
spde <- inla.spde2.matern(mesh=mesh,alpha=2)
s.index <- inla.spde.make.index(name="spatial.field", n.spde=spde$n.spde)

A_y <- inla.spde.make.A(mesh=mesh, loc=as.matrix(sf.plots %>% as.data.frame %>% select(X,Y)))
A_pred <- inla.spde.make.A(mesh=mesh, loc=as.matrix(sf.grid %>% as.data.frame %>% select(X,Y)))

stack.est <- inla.stack(data=list(y = sf.plots[["Struikhei"]]),
                        A=list(A_y), effects=list(c(s.index, list(intercept=1))), 
                        tag="est")
stack.pred <- inla.stack(data=list(y=NA), 
                         A=list(A_pred), effects=list(c(s.index, list(intercept=1))),
                         tag="pred") 
joint.stack <- inla.stack(stack.est, stack.pred)

formula <- y ~ -1 + intercept + f(spatial.field, model=spde)

output.stack <- inla(formula, family="binomial", 
               data=inla.stack.data(joint.stack, spde=spde),
               control.predictor=list(A=inla.stack.A(joint.stack), compute=TRUE, link=1))

# Save extrapolated Struikhei as indicator of pH
plot.indicator <- output.stack$summary.linear.predictor[inla.stack.index(joint.stack, tag="est")$data,"mean"]
grid.indicator <- output.stack$summary.linear.predictor[inla.stack.index(joint.stack, tag="pred")$data,"mean"]
sf.plots$pHindicator <- (plot.indicator - mean(grid.indicator))/sd(grid.indicator)
sf.grid$pHindicator <- (grid.indicator - mean(grid.indicator))/sd(grid.indicator)
```

Plot observed vs. predicted common heather prevalence

```{r}
# Calculate observed Strukhei at each grid point
results <- sf.grid %>% st_set_geometry("geometry.1") %>% 
  st_join(sf.plots%>%select(`Struikhei`), st_contains) %>% 
  group_by(geometry.1) %>% summarize(Struikhei=mean(`Struikhei`, na.rm=T))
# Save predicted Strukhei at each grid point
index.pred.response <- inla.stack.index(joint.stack, tag="pred")$data
results$Struikhei.mean <- output.stack$summary.fitted.values[index.pred.response, "mean"]
results$Struikhei.sd <- output.stack$summary.fitted.values[index.pred.response,"sd"]

plot1 <- ggplot(NL) + geom_sf() +
	geom_sf(aes(geometry=geometry.1, fill=Struikhei), data=results) + 
  scale_fill_gradient(low="lightyellow", high="green", limits=c(0,1))+
	ggtitle('Observed Struikhei') + theme(legend.position="bottom")

plot2 <- ggplot(NL) + geom_sf() +
	geom_sf(aes(geometry=geometry.1, fill=Struikhei.mean), data=results) + 
  scale_fill_gradient(low="lightyellow", high="green", limits=c(0,1))+
	ggtitle('Predicted Struikhei') + theme(legend.position="bottom")

plot_grid(plot1, plot2, nrow = 1)
ggsave('results_paper/Struikhei.png', width=2000, heigh=1400, units='px')
```

## Spatial interpolation of pH

Illustrate the mesh and pH measurements:

```{r}
plot(NL, col="grey", main='Gaussian Markov Random Field')
plot(mesh, asp=1, add=T)
rbPal <- colorRampPalette(c('blue','red'))
col <- rbPal(10)[as.numeric(cut(sf.plots[["pH-H20"]],breaks = 10))]
plot(sf.plots["pH-H20"], pch=21, bg=col, pal=rbPal, cex=0.5, add=T)
legend(-200,150, text.font=4, cex=0.6, legend=c("Low", "High"), col=c("blue", "red"), pch=21)#16000, 614000
```

Interpolate pH in the Netherlands with kriging using soil type, landuse, and indicator plant (common heather) as covariates

```{r}
A_y <- inla.spde.make.A(mesh=mesh, loc=as.matrix(sf.plots %>% as.data.frame %>% select(X,Y)))
A_pred <- inla.spde.make.A(mesh=mesh, loc=as.matrix(sf.grid %>% as.data.frame %>% select(X,Y)))

stack.est <- inla.stack(data=list(y = sf.plots[["pH-H20"]]),
                        A=list(A_y,1,1,1), effects=list(c(s.index, list(intercept=1)),
                                                  list(OMSCHRIJVI = sf.plots$OMSCHRIJVI),
                                                  list(bodemgebruik = sf.plots$bodemgebruik),
                                                  list(pHindicator = sf.plots$pHindicator)), 
                        tag="est")
stack.pred <- inla.stack(data=list(y=NA), 
                         A=list(A_pred,1,1,1), effects=list(c(s.index, list(intercept=1)),
                                                          list(OMSCHRIJVI = sf.grid$OMSCHRIJVI),
                                                          list(bodemgebruik = sf.grid$bodemgebruik),
                                                          list(pHindicator = sf.grid$pHindicator)),
                         tag="pred") 
joint.stack <- inla.stack(stack.est, stack.pred)

formula <- y ~ -1 + intercept + OMSCHRIJVI + bodemgebruik + pHindicator +f(spatial.field, model=spde) #+bodemgebruik

output.stack <- inla(formula, family="gaussian", 
               data=inla.stack.data(joint.stack, spde=spde),
               #control.fixed = list(expand.factor.strategy = "inla"),
               control.predictor=list(A=inla.stack.A(joint.stack), compute=TRUE))

output.field <- inla.spde2.result(inla=output.stack, name="spatial.field", spde=spde, do.transf=TRUE)
```

Illustrate the model parameters: intercept, measurement error, spatial variance and spatial range

```{r}
### Plot posterior distributions
par(mfrow=c(2,2))
# Plot for beta0
xrange <- range(output.stack$marginals.fix[[1]][,1])
yrange <- range(output.stack$marginals.fix[[1]][,2])
plot(output.stack$marginals.fix[[1]], type='l', xlim=xrange, ylim=yrange, xlab=expression(b[0]),ylab="")

# Plot for the variance sigma2e
s2e.marg <- inla.tmarginal(function(x) 1/x, output.stack$marginals.hy[[1]])
xrange <- range(s2e.marg[,1])
yrange <- range(s2e.marg[,2])
plot.default(s2e.marg, type='l', xlim=xrange, ylim=yrange, xlab=expression(sigma[e]^2),ylab="")

# Plot for the variance sigma2
xrange <- range(output.field$marginals.variance.nominal[[1]][,1])
yrange <- range(output.field$marginals.variance.nominal[[1]][,2])
plot(output.field$marginals.variance.nominal[[1]], type='l', xlim=xrange, ylim=yrange, xlab=expression(sigma^2),ylab="")

#Plot for the range
xrange <- range(output.field$marginals.range.nominal[[1]][,1])
yrange <- range(output.field$marginals.range.nominal[[1]][,2])
plot(output.field$marginals.range.nominal[[1]], type='l', xlim=xrange, ylim=yrange, xlab=expression(r),ylab="")

```

Plot observed vs. predicted pH

```{r}
# Calculate observed pH at each grid point
results <- sf.grid %>% st_set_geometry("geometry.1") %>% 
  st_join(sf.plots%>%select(`pH-H20`), st_contains) %>% 
  group_by(geometry.1) %>% summarize(pH=mean(`pH-H20`, na.rm=T))
# Save predicted pH at each grid point
index.pred.response <- inla.stack.index(joint.stack, tag="pred")$data
results$pH.mean <- output.stack$summary.fitted.values[index.pred.response,"mean"]
results$pH.sd <- output.stack$summary.fitted.values[index.pred.response,"sd"]

# Plot (ggplot)
plot1 <- ggplot(NL) + geom_sf() + geom_sf(aes(geometry=geometry.1, fill=pH), data=results) + 
  scale_fill_gradient2(low='red', midpoint=7, high='blue') + ggtitle('Observed pH') + theme(legend.position="bottom")
plot2 <- ggplot(NL) + geom_sf() + geom_sf(aes(geometry=geometry.1, fill=pH.mean), data=results) + 
  scale_fill_gradient2(low='red', midpoint=7, high='blue') + ggtitle('Predicted pH') + theme(legend.position="bottom")
#plot_grid(plot1, plot2, nrow = 1) + theme(plot.background = element_rect(fill = "white", colour = NA)) #, plot3
#ggsave('results_paper/pH.png', width=2000, heigh=1400, units='px')

plot.all <- ggdraw() +
  draw_plot(plot1, x = 0, y = 0, width = 0.50, height = 0.95) +
  draw_plot(plot2, x = 0.50, y = 0, width = 0.50, height = 0.95) +
  draw_plot_label(label = c("A", "B"), size = 15, x = c(0.01, 0.50), y = c(0.99,0.99)) +
  theme(plot.background = element_rect(fill = "white", colour = NA))
plot.all
ggsave('results_paper/pH.png', width=2000, heigh=1400, units='px')
```

Same plot with a third panel visualizing uncertainty of interpolated pH

```{r}
plot3 <- ggplot(NL) + geom_sf() + geom_sf(aes(geometry=geometry.1, fill=pH.sd), data=results) + 
  scale_fill_gradient(low='white', high='red', limits=c(0,1.11)) + ggtitle('Uncertainty pH') + theme(legend.position="bottom")
plot_grid(plot1, plot2, plot3, nrow = 1) + theme(plot.background = element_rect(fill = "white", colour = NA)) #, plot3
ggsave('results_paper/pH2.png', width=3000, heigh=1400, units='px')
```

Save interpolated pH for modelling

```{r}
index.pred.response <- inla.stack.index(joint.stack, tag="est")$data
sf.plots$pH <- output.stack$summary.fitted.values[index.pred.response,"mean"]
index.pred.response <- inla.stack.index(joint.stack, tag="pred")$data
sf.grid$pH <- output.stack$summary.fitted.values[index.pred.response,"mean"]
```

Histogram of pH in grid cells vs. vegetation plots

```{r}
f <- rbind(data.frame(pH = sf.grid$pH, Data='Grid'),
           data.frame(pH = sf.plots$pH, Data='Plots'))

ggplot(f, aes(x=pH, fill=Data)) +
  geom_histogram(aes(y=after_stat(density)), bins=50, position="dodge") +
  scale_fill_manual(values=c('#999999','#E69F00'))

```

## Process data

Marginal distribution statistics:

```{r}
n <- sum(sf.plots$is_fieldvisit)
sf.plots %>% as.data.frame %>% transmute(
  is_fieldvisit, has_abiotics, Dagnummer = Dagnummer, 
  pH = pH, ORG_STOF = log(ORG_STOF), CN = log(`C/N`), 
  N_totaal = log(N_totaal), P_totaal = log(P_totaal),
  K_CaCl2 = log(ifelse(K_CaCl2<=0, 0.1, K_CaCl2))) %>%
  summarize(across(all_of(c('Dagnummer', 'pH', 'ORG_STOF', 'CN','N_totaal', 'P_totaal', 'K_CaCl2')), 
                   list(mean=function(x) mean(x,na.rm=T), sd=function(x) sd(x,na.rm=T),
                        available = function(x) sum(!is.na(x) & is_fieldvisit)/n))) %>% round(2)
```
Z-score abiotic variables and clip extreme outliers to ensure equal impact of priors

```{r}
to_zscore <- function(x) (x - mean(x, na.rm=T))/sd(x, na.rm=T)
zclip <- function(x) ifelse(x > 3, 3, ifelse(x < -3, -3, x))

pH.mean = 7 #mean(sf.plots$pH)
pH.sd = 1 #sd(sf.plots$pH)

sf.plots <- sf.plots %>% mutate(
 Pq_grootte=replace_na(Pq_grootte, 25),
 Dagnummer = replace_na(Dagnummer, 200),
 Dagnummer.z = (Dagnummer-200)/30.44, 
 pH.z = zclip((pH - pH.mean)/pH.sd), 
 ORG_STOF.z = zclip(to_zscore(log(ORG_STOF))), 
 CN.z = zclip(to_zscore(log(`C/N`))),
 N_totaal.z = zclip(to_zscore(log(N_totaal))), 
 P_totaal.z = zclip(to_zscore(log(P_totaal))), 
 K_CaCl2.z = zclip(to_zscore(log(ifelse(K_CaCl2<=0, 0.1, K_CaCl2)))) 
)

sf.grid <- sf.grid %>% mutate(
 Dagnummer=200, Pq_grootte=25,
 Dagnummer.z = (Dagnummer-200)/30.44, 
 pH.z =  zclip((pH - pH.mean)/pH.sd)
)

```

Landuse coding for the spline

```{r}
code_terrain <- function (x) switch(x, "Bos"=-4, "Droog natuurlijk terrein"=-3, "Nat natuurlijk terrein"=-2, "Water"=-1,
                                     "Recreatie"=0, "Spoorterrein"=1, "Hoofdweg"=2,  "Bebouwd"=3, "Landbouw en overig agrarisch"=4)
sf.plots$terrain <- sapply(sf.plots$bodemgebruik %>% as.character, code_terrain)
sf.grid$terrain <- sapply(sf.grid$bodemgebruik %>% as.character, code_terrain)
```

Matching coordinate matrix

```{r}
xy.plots <- sf.plots %>% as.data.frame %>% select(X,Y)
xy.grid <- sf.grid %>% as.data.frame %>% select(X,Y)
```

For the cluster use these pre-processed files

```{r}
sf.plots %>% saveRDS("sf.plots.RDS")
sf.grid %>% saveRDS("sf.grid.RDS")
```

## Experiments: fit both models to data

Uniform mesh for the spatial component

```{r}
step <- 10000 / 1000
domain <- inla.nonconvex.hull(as.matrix(xy.plots), convex=-0.04)
mesh <- inla.mesh.2d(boundary=domain, max.edge=c(step*2, step*4), cutoff=step)
```

```{r}
spde <- inla.spde2.matern(mesh=mesh,alpha=2)##print(spde$n.spde)
s.index <- inla.spde.make.index(name="spatial.field", n.spde=spde$n.spde)

A_y <- inla.spde.make.A(mesh=mesh, loc=as.matrix(xy.plots))
A_pred <- inla.spde.make.A(mesh=mesh, loc=as.matrix(xy.grid))
```

### Two-stage model

Predict missing abiotic variables

```{r}
terms.abiotics <- list()
i <- 1
for (variable in c("ORG_STOF.z", "CN.z", "N_totaal.z", "P_totaal.z", "K_CaCl2.z")) {
  print(variable)

  stack.est <- inla.stack(data=list(y = sf.plots[[variable]]),
                          A=list(A_y), effects=list(c(s.index, list(intercept=1))), 
                          tag="est")
  stack.pred <- inla.stack(data=list(y=NA), 
                           A=list(A_pred), effects=list(c(s.index, list(intercept=1))),
                           tag="pred") 
  join.stack <- inla.stack(stack.est, stack.pred)

  formula <- y ~ -1 + intercept + f(spatial.field, model=spde)

  output <- inla(formula, family="gaussian", 
                      data=inla.stack.data(join.stack, spde=spde),
                      control.predictor=list(A=inla.stack.A(join.stack), compute=TRUE, link=1))

  # predictions at field visits
  index.est <- inla.stack.index(join.stack, tag="est")$data
  variable.est <- output$summary.fitted.values[index.est,]
  colnames(variable.est) <- sprintf("%s.%s", variable, colnames(variable.est))
  sf.plots[colnames(variable.est)] <- variable.est

  # predictions at grid
  index.pred <- inla.stack.index(join.stack, tag="pred")$data
  variable.pred <- output$summary.fitted.values[index.pred,]
  colnames(variable.pred) <- sprintf("%s.%s", variable, colnames(variable.pred))
  sf.grid[colnames(variable.pred)] <- variable.pred

  # transform posterior marginal of precision into variance and calculate statistics
  i_field <- inla.spde2.result(output, name="spatial.field", spde)
  var.marginal <- data.frame(inla.zmarginal(inla.tmarginal(function(x) log(1/x), output$marginals.hyperpar[[1]]), silent=T)) %>%
    rename(mean=mean, sd=sd, `0.025quant`=quant0.025, `0.25quant`=quant0.25, `0.5quant`=quant0.5, `0.75quant`=quant0.75, `0.975quant`=quant0.975)
  # predictions
  cols <- sprintf(c("%s.mean", "%s.sd", "%s.0.025quant", "%s.0.5quant", "%s.0.975quant"), variable)
  i.pred <- sf.grid %>% as.data.frame %>% select(all_of(cols))
  colnames(i.pred) <- c("mean", "sd", "0.025quant", "0.5quant", "0.975quant")

  # Abiotic factor regression terms and predictions
  terms.abiotics[[variable]] <- rbind(
  	# Intercept
  	output$summary.fixed %>% 
  		mutate(ID=row.names(output$summary.fixed)) %>% remove_rownames %>% mutate(term="fixed") %>% 
  		select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`),
    # Precision
  	var.marginal %>% 
  		mutate(ID="log.variance.eps") %>% remove_rownames %>% mutate(term="spatial") %>% 
  		select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`),
  	# Spatial component
  	bind_rows(list(log.variance=i_field$summary.log.variance.nominal, 
  	               log.range=i_field$summary.log.range.nominal), .id="ID") %>% 
  	  remove_rownames %>% mutate(term="spatial") %>% 
  		select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`),
    # Predictions
    cbind(sf.grid %>% as.data.frame %>% mutate(term="Predict", ID=Plot), i.pred) %>% 
    select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`)
  	)
  i <- i + 1
}

terms.abiotics <- bind_rows(terms.abiotics, .id="model") %>% 
  mutate(species="", meta="Certain") %>%
  select(species, meta, model, term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`)

write.table(terms.abiotics, "results_certain/abiotics.csv", sep = ",", col.names = T, row.names=F, append=F)
```

Visualization of predicted abiotic variables

```{r}
for (var in c("ORG_STOF.z", "CN.z", "N_totaal.z", "P_totaal.z", "K_CaCl2.z")) {
  var.mean <- sprintf("%s.mean", var)
  var.sd <- sprintf("%s.sd", var)

  # Take predicted abiotic in the grid and observed in the plot
  results <- sf.grid %>% select(geometry.1, var.mean,  var.sd) %>% st_set_geometry("geometry.1") %>%
    st_join(sf.plots %>% select(geometry, var), st_contains) %>% group_by(geometry.1) %>% summarize(
      y.mean=mean(get(var.mean), na.rm=T), 
      y.sd=mean(get(var.sd), na.rm=T), 
      y=mean(get(var), na.rm=T)
      )

  # Plot (ggplot)
  plot1 <- ggplot(NL) + geom_sf() + geom_sf(aes(geometry=geometry.1, fill=y), data=results) + 
    scale_fill_gradient2(low='blue', midpoint=0, high='red', limits=c(-3,3)) + 
    ggtitle(sprintf('Observed %s', var)) + theme(legend.position="bottom")
  plot2 <- ggplot(NL) + geom_sf() + geom_sf(aes(geometry=geometry.1, fill=y.mean), data=results) + 
    scale_fill_gradient2(low='blue', midpoint=0, high='red', limits=c(-3,3)) + 
    ggtitle(sprintf('Predicted %s', var)) + theme(legend.position="bottom")
  plot3 <- ggplot(NL) + geom_sf() + geom_sf(aes(geometry=geometry.1, fill=y.sd), data=results) + 
    scale_fill_gradient(low='white', high='red', limits=c(0,1)) + #scale_fill_gradient(low='darkblue', high='yellow', limits=c(0,1)) + 
    ggtitle(sprintf('Uncertainty %s', var)) + theme(legend.position="bottom")
  plot_grid(plot1, plot2, plot3, nrow = 1)
}
```

Fit the two-stage model using predicted abiotic variables, save model parameters and predictions

```{r}

for (species.take in species) { #c("Gestreepte witbol")
  print(species.take)

  start <- Sys.time()
  stk_y <- inla.stack(data = list(y=sf.plots[[species.take]]),
                A = list(A_y, 1, 1, 1, 1, 1, 1, 1, 1), 
                effects = list(c(list(u.field=1:spde$n.spde), list(intercept=1)), 
                               list(terrain=sf.plots$terrain), 
                               list(Dagnummer.z=sf.plots$Dagnummer.z), 
                               list(pH.z=sf.plots$pH.z), 
                               list(ORG_STOF.z=sf.plots$ORG_STOF.z.mean),
                               list(CN.z=sf.plots$CN.z.mean), 
                               list(N_totaal.z=sf.plots$N_totaal.z.mean), 
                               list(P_totaal.z=sf.plots$P_totaal.z.mean), 
                               list(K_CaCl2.z=sf.plots$K_CaCl2.z.mean)
                               ),
                tag="est.y")

  n <- nrow(sf.grid)
  stk_pred <- inla.stack(data=list(y=rep(NA, n)),  
                          effects=list(c(list(u.field=1:spde$n.spde), list(intercept=1)),
                                       list(terrain=sf.grid$terrain), 
                                       list(Dagnummer.z=sf.grid$Dagnummer.z), 
                                       list(pH.z=sf.grid$pH.z), 
                                       list(ORG_STOF.z=sf.grid$ORG_STOF.z.mean),
                                       list(CN.z=sf.grid$CN.z.mean), 
                                       list(N_totaal.z=sf.grid$N_totaal.z.mean), 
                                       list(P_totaal.z=sf.grid$P_totaal.z.mean), 
                                       list(K_CaCl2.z=sf.grid$K_CaCl2.z.mean)
                               ),
                          A=list(A_pred, 1, 1, 1, 1, 1, 1, 1, 1),
                          tag="est.pred")

  stk <- inla.stack(stk_y, stk_pred)

  formula <- y ~ -1 + intercept + f(inla.group(terrain), model = "rw2") + 
    f(inla.group(Dagnummer.z), model = "rw2") +  f(inla.group(pH.z), model = "rw2") +
    CN.z + ORG_STOF.z + N_totaal.z + P_totaal.z + K_CaCl2.z + f(u.field, model=spde)

  output <-  tryCatch(
        {inla(formula, num.threads = 1,
                  data=inla.stack.data(stk, spde=spde), family="binomial",
                  control.predictor=list(A=inla.stack.A(stk), compute=TRUE, link=1), #
                  control.inla = list(strategy = "adaptive", int.strategy = "eb")) #int.strategy = "eb"
        }, error = function(cond) 
            {message(cond)
            return(NULL)}
    )

  # Model results
  if (!is.null(output)) {
    u_field <- inla.spde2.result(output, name="u.field", spde)
    index.pred <- inla.stack.index(stk, tag="est.pred")$data

    # SDM regression terms and predictions
    terms.sdm <- rbind(
      # Splines for day number, pH, terrain
    	bind_rows(output$summary.random, .id="term") %>% 
    			filter(term %in% c("inla.group(Dagnummer.z)", "inla.group(pH.z)", "inla.group(terrain)")) %>% 
    	    mutate(term=str_replace_all(term, c("inla.group\\("="", "\\)"=""))) %>% 
    			select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`),
    	# Linear terms for CN, ORG_STOF, N_totaal, P_totaal, K_CaCl2
    	output$summary.fixed %>% 
    		mutate(ID=row.names(output$summary.fixed)) %>% remove_rownames %>% mutate(term="fixed") %>% 
    		select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`),
    	# Spatial component
    	bind_rows(list(log.variance=u_field$summary.log.variance.nominal, 
    	               log.range=u_field$summary.log.range.nominal), .id="ID") %>% 
    	  remove_rownames %>% mutate(term="spatial") %>% 
    		select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`),
      # Predictions
    	cbind(sf.grid %>% as.data.frame %>% mutate(term="Predict", ID=Plot), 
                       output$summary.fitted.values[index.pred,]) %>% 
      select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`),
      # Predictions (log-odds scale)
    	cbind(sf.grid %>% as.data.frame %>% mutate(term="APredict", ID=Plot), 
                       output$summary.linear.predictor[index.pred,]) %>% 
      select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`)
    ) %>% 
      mutate(species=species.take, meta="Certain", model = "SDM") %>%
      select(species, meta, model, term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`)

    # Save
    terms <- rbind(terms.sdm)#, terms.abiotics %>% mutate(species=species.take)
    fn <- sprintf('results_certain/%s.csv', species.take)
    write.table(terms, fn, sep = ",", col.names = T, row.names=F, append=F)
    #write.table(terms, "results_certain.csv", sep = ",", col.names = !file.exists("results_certain.csv"), row.names=F, append=T)
  }

  end <- Sys.time()
  print(end-start)
}
```

### Joint model

Construct a separate model for each abiotic variable:

```{r}
variables <- c("ORG_STOF.z", "CN.z", "N_totaal.z", "P_totaal.z", "K_CaCl2.z")

i <- 1
stacks <- list()
for (var in variables) {
  # Observations of the variable
  sf.var <- sf.plots %>% filter(!is.na(get(var)))
  A_var <- inla.spde.make.A(mesh=mesh, loc=as.matrix(sf.var %>% as.data.frame %>% select(X,Y)))
  # One output for every estimated variable (+ species occurrence)
  y <- matrix(nrow=nrow(sf.var), ncol=length(variables) + 1)
  y.0 <- matrix(nrow=nrow(sf.grid), ncol=length(variables) + 1)
  y[,i] <- sf.var[[var]]
  # Stack
  effects <- list()
  effects[[sprintf("%s.i.field", var)]] = 1:spde$n.spde
  effects[[sprintf("%s.intercept", var)]] = 1 
  stacks[[var]] <- inla.stack(data=list(y=y), effects=list(effects), A=list(A_var), tag=sprintf("est.%s", var))
  stacks[[sprintf("%s.pred", var)]] <- inla.stack(data=list(y=y.0), effects=list(effects), A=list(A_pred), tag=sprintf("est.pred.%s", var))
  i <- i + 1
}
```

Fit the joint model using models for abiotic variables, save model parameters and predictions

```{r}

for (species.take in species) { #c("Gestreepte witbol")
  print(species.take)

  start <- Sys.time()
  # Species occurrence
  y <- matrix(nrow=nrow(sf.plots), ncol=length(variables) + 1)
  y[,i] <- sf.plots[[species.take]]
  # Stack
  stk_y <- inla.stack(data=list(y=y),  
                          effects=list(c(list(ORG_STOF.z.field=1:spde$n.spde, 
                                              CN.z.field=1:spde$n.spde,
                                              N_totaal.z.field=1:spde$n.spde, 
                                              P_totaal.z.field=1:spde$n.spde, 
                                              K_CaCl2.z.field=1:spde$n.spde,
                                              u.field=1:spde$n.spde), list(intercept=1)),
                                       list(terrain=sf.plots$terrain), 
                                       list(Dagnummer.z=sf.plots$Dagnummer.z), 
                                       list(pH.z=sf.plots$pH.z)),
                          A=list(A_y, 1, 1, 1),
                          tag="est.y")
  # Predicted over grid
  y <- matrix(nrow=nrow(sf.grid), ncol=length(variables) + 1)
  stk_pred <- inla.stack(data=list(y=y),  
                          effects=list(c(list(ORG_STOF.z.field=1:spde$n.spde, 
                                              CN.z.field=1:spde$n.spde,
                                              N_totaal.z.field=1:spde$n.spde, 
                                              P_totaal.z.field=1:spde$n.spde, 
                                              K_CaCl2.z.field=1:spde$n.spde,
                                              u.field=1:spde$n.spde), list(intercept=1)),
                                       list(terrain=sf.grid$terrain), 
                                       list(Dagnummer.z=sf.grid$Dagnummer.z), 
                                       list(pH.z=sf.grid$pH.z)),
                          A=list(A_pred, 1, 1, 1),
                          tag="est.pred")

  stk <- inla.stack(stacks[["ORG_STOF.z"]], stacks[["ORG_STOF.z.pred"]],
                    stacks[["CN.z"]], stacks[["CN.z.pred"]], 
                    stacks[["N_totaal.z"]], stacks[["N_totaal.z.pred"]],
                    stacks[["P_totaal.z"]], stacks[["P_totaal.z.pred"]],
                    stacks[["K_CaCl2.z"]], stacks[["K_CaCl2.z.pred"]],
                    stk_y, stk_pred)

  formula <- y ~ -1 + intercept + ORG_STOF.z.intercept + CN.z.intercept + N_totaal.z.intercept + P_totaal.z.intercept + K_CaCl2.z.intercept +
    f(inla.group(terrain), model = "rw2") + f(inla.group(Dagnummer.z), model = "rw2") +  f(inla.group(pH.z), model = "rw2") + 
    f(ORG_STOF.z.i.field, model=spde) + f(ORG_STOF.z.field, copy="ORG_STOF.z.i.field", fixed=FALSE, hyper=list(theta=list(param=c(0, 0.001)))) +
    f(CN.z.i.field, model=spde) + f(CN.z.field, copy="CN.z.i.field", fixed=FALSE, hyper=list(theta=list(param=c(0, 0.001)))) +
    f(N_totaal.z.i.field, model=spde) + f(N_totaal.z.field, copy="N_totaal.z.i.field", fixed=FALSE, hyper=list(theta=list(param=c(0, 0.001))))  +
    f(P_totaal.z.i.field, model=spde) + f(P_totaal.z.field, copy="P_totaal.z.i.field", fixed=FALSE, hyper=list(theta=list(param=c(0, 0.001))))  +
    f(K_CaCl2.z.i.field, model=spde) + f(K_CaCl2.z.field, copy="K_CaCl2.z.i.field", fixed=FALSE, hyper=list(theta=list(param=c(0, 0.001)))) +
    f(u.field, model=spde)

  families <- c(rep("gaussian", length(variables)), "binomial")

  output <-  tryCatch(
        {inla(formula, family=families, num.threads=1,
                 data=inla.stack.data(stk), 
                 control.predictor=list(A=inla.stack.A(stk), compute=TRUE, link=i),
                 control.inla = list(strategy = "adaptive", int.strategy = "eb"))#int.strategy = "eb"
        }, error = function(cond) 
            {message(cond)
            return(NULL)}
    )

  # Model results
  if (!is.null(output)) {
    # random field and index in stack
    u_field <- inla.spde2.result(output, name="u.field", spde)
    index.pred <- inla.stack.index(stk, tag="est.pred")$data

    # SDM regression terms and predictions
    terms.sdm <- rbind(
      # Splines for day number, pH, terrain
    	bind_rows(output$summary.random, .id="term") %>% 
    			filter(term %in% c("inla.group(Dagnummer.z)", "inla.group(pH.z)", "inla.group(terrain)")) %>% 
    	    mutate(term=str_replace_all(term, c("inla.group\\("="", "\\)"=""))) %>% 
    			select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`),
    	# Intercept
    	output$summary.fixed %>% 
    		mutate(ID=row.names(output$summary.fixed)) %>% remove_rownames %>% mutate(term="fixed") %>% 
    		filter(ID  == "intercept") %>%
    		select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`),
    	# Linear terms for CN, ORG_STOF, N_totaal, P_totaal, K_CaCl2
    	output$summary.hyperpar %>% 
    		mutate(ID=row.names(output$summary.hyperpar)) %>% remove_rownames %>% mutate(term="fixed") %>% 
    		filter(ID %>% startsWith("Beta")) %>%
    	  mutate(ID=str_replace_all(ID, c("Beta for "="", ".field"=""))) %>%
    		select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`),
    	# Spatial component
    	bind_rows(list(log.variance=u_field$summary.log.variance.nominal, 
    	               log.range=u_field$summary.log.range.nominal), .id="ID") %>% 
    	  remove_rownames %>% mutate(term="spatial") %>% 
    		select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`),
      # Predictions
    	cbind(sf.grid %>% as.data.frame %>% mutate(term="Predict", ID=Plot), 
                       output$summary.fitted.values[index.pred,]) %>% 
      select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`),
      # Predictions (log-odds scale)
    	cbind(sf.grid %>% as.data.frame %>% mutate(term="APredict", ID=Plot), 
                       output$summary.linear.predictor[index.pred,]) %>% 
      select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`)
    ) %>% 
      mutate(species=species.take, meta="Uncertain", model = "SDM") %>%
      select(species, meta, model, term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`)

    j <- 1
    terms.abiotics <- list()
    for (var in variables) {
      ## transform posterior marginal of precision into variance and calculate statistics
      i_field <- inla.spde2.result(output, name=sprintf("%s.i.field", var), spde)
      var.marginal <- data.frame(inla.zmarginal(inla.tmarginal(function(x) log(1/x), output$marginals.hyperpar[[j]]), silent=T)) %>%
        rename(mean=mean, sd=sd, `0.025quant`=quant0.025, `0.25quant`=quant0.25, `0.5quant`=quant0.5, `0.75quant`=quant0.75, `0.975quant`=quant0.975)
      # predictions
      index.pred <- inla.stack.index(stk, tag=sprintf("est.pred.%s", var))$data
      i.pred <- output$summary.linear.predictor[index.pred,] %>% select(mean, sd, `0.025quant`, `0.5quant`,`0.975quant`)

      # Abiotic factor regression terms and predictions
      terms.abiotics[[var]] <- rbind(
      	# Intercept
      	output$summary.fixed %>% 
      		mutate(ID=row.names(output$summary.fixed)) %>% remove_rownames %>% mutate(term="fixed") %>% 
      		filter(ID  == sprintf("%s.intercept", var)) %>% mutate(ID = "intercept") %>% 
      		select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`),
        # Precision
      	var.marginal %>% 
      		mutate(ID="log.variance.eps") %>% remove_rownames %>% mutate(term="spatial") %>% 
      		select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`),
      	# Spatial component
      	bind_rows(list(log.variance=i_field$summary.log.variance.nominal, 
      	               log.range=i_field$summary.log.range.nominal), .id="ID") %>% 
      	  remove_rownames %>% mutate(term="spatial") %>% 
      		select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`),
        # Predictions
        cbind(sf.grid %>% as.data.frame %>% mutate(term="Predict", ID=Plot), i.pred) %>% 
        select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`)
      )	
      j <- j + 1
    }
    terms.abiotics <- bind_rows(terms.abiotics, .id="model") %>% 
      mutate(species=species.take, meta="Uncertain") %>%
      select(species, meta, model, term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`)

    # Save
    terms <- rbind(terms.sdm, terms.abiotics)
    fn <- sprintf('results_uncertain/%s.csv', species.take)
    write.table(terms, fn, sep = ",", col.names = T, row.names=F, append=F)
    #write.table(terms, "results_uncertain.csv", sep = ",", col.names = !file.exists("results_uncertain.csv"), row.names=F, append=T)

  }

  end <- Sys.time()
  print(end-start)
}
```

## Experiments: training and validation

### Provinces

Two-stage model (certain abiotics), save predictions and prevalences

```{r}
provincies <- sort(unique(sf.plots$naam))
for (species.take in species) { 
  print(species.take)
  for (provincie in provincies) {#c("Utrecht")
    print(provincie)
    sf.train = sf.plots %>% filter(naam != provincie) 
    sf.test = sf.plots %>% filter(naam == provincie) 
    A_train <- inla.spde.make.A(mesh=mesh, loc=as.matrix(sf.train %>% as.data.frame %>% select(X,Y)))
    A_test <- inla.spde.make.A(mesh=mesh, loc=as.matrix(sf.test %>% as.data.frame %>% select(X,Y)))

    # Fit
    stk_y <- inla.stack(data = list(y=sf.train[[species.take]]),
                  A = list(A_train, 1, 1, 1, 1, 1, 1, 1, 1), 
                  effects = list(c(list(u.field=1:spde$n.spde), list(intercept=1)), 
                                 list(terrain=sf.train$terrain), 
                                 list(Dagnummer.z=sf.train$Dagnummer.z), 
                                 list(pH.z=sf.train$pH.z), 
                                 list(ORG_STOF.z=sf.train$ORG_STOF.z.mean),
                                 list(CN.z=sf.train$CN.z.mean), 
                                 list(N_totaal.z=sf.train$N_totaal.z.mean), 
                                 list(P_totaal.z=sf.train$P_totaal.z.mean), 
                                 list(K_CaCl2.z=sf.train$K_CaCl2.z.mean)
                                 ),
                  tag="est.y")

    stk_pred <- inla.stack(data=list(y=rep(NA, nrow(sf.test))),  
                            effects=list(c(list(u.field=1:spde$n.spde), list(intercept=1)),
                                         list(terrain=sf.test$terrain), 
                                         list(Dagnummer.z=sf.test$Dagnummer.z), 
                                         list(pH.z=sf.test$pH.z), 
                                         list(ORG_STOF.z=sf.test$ORG_STOF.z.mean),
                                         list(CN.z=sf.test$CN.z.mean), 
                                         list(N_totaal.z=sf.test$N_totaal.z.mean), 
                                         list(P_totaal.z=sf.test$P_totaal.z.mean), 
                                         list(K_CaCl2.z=sf.test$K_CaCl2.z.mean)
                                 ),
                            A=list(A_test, 1, 1, 1, 1, 1, 1, 1, 1),
                            tag="est.pred")

    stk <- inla.stack(stk_y, stk_pred)

    formula <- y ~ -1 + intercept + f(inla.group(terrain), model = "rw2") + 
      f(inla.group(Dagnummer.z), model = "rw2") +  f(inla.group(pH.z), model = "rw2") +
      CN.z + ORG_STOF.z + N_totaal.z + P_totaal.z + K_CaCl2.z + f(u.field, model=spde)

    index.pred <- inla.stack.index(stk, tag="est.pred")$data

    output <-  tryCatch(
        {inla(formula, num.threads=1,
                    data=inla.stack.data(stk, spde=spde), family="binomial",
                    control.predictor=list(A=inla.stack.A(stk), compute=TRUE, link=1), 
                    control.inla = list(strategy = "adaptive", int.strategy = "eb"), control.compute=list(config = TRUE)) #int.strategy = "eb"
        }, error = function(cond) 
            {message(cond)
            return(NULL)}
    )

    # First model results
    if (!is.null(output)) {
      # get inla posterior sample
      s1 <- inla.posterior.sample(400, output) 
      linear_predictor <- inla.posterior.sample.eval("APredictor", s1)[index.pred,]
      P = 1/(1+exp(-linear_predictor))
      sample_mean <- function(p) mean(rbinom(length(p), 1, p))
      area_prevalences <- apply(P, 2, sample_mean) #colMeans(1/(1+exp(-linear_predictor)))

      # Predicted vs observed
      predicted <- output$summary.fitted.values[index.pred,c("mean", "sd", "0.025quant", "0.975quant")]
      observed <- sf.test[[species.take]]

      predictions <- cbind(predicted, observed) %>% mutate(meta = "Certain", model="SDM", species=species.take, naam = provincie)
      prevalences <- data.frame(area_prevalences, meta = "Certain", model="SDM", species=species.take, naam = provincie)
      #write.table(predictions, "predictions_certain.csv", sep = ",", col.names = !file.exists("predictions_certain.csv"), row.names=F, append = T)
      #write.table(prevalences, "prevalences_certain.csv", sep = ",", col.names = !file.exists("prevalences_certain.csv"), row.names=F, append = T)
      fn.1 <- sprintf('predictions_certain/%s_%s.csv', species.take, provincie)
      fn.2 <- sprintf('prevalences_certain/%s_%s.csv', species.take, provincie)
      write.table(predictions, fn.1, sep = ",", col.names = T, row.names=F, append=F)
      write.table(prevalences, fn.2, sep = ",", col.names = T, row.names=F, append=F)
    }

  }
}
```

Joint model (uncertain abiotics), save predictions and prevalences

```{r}
provincies <- sort(unique(sf.plots$naam))
for (species.take in species) { 
  print(species.take)
  for (provincie in provincies) {#c("Utrecht")
    print(provincie)
    sf.train = sf.plots %>% filter(naam != provincie) 
    sf.test = sf.plots %>% filter(naam == provincie) 
    A_train <- inla.spde.make.A(mesh=mesh, loc=as.matrix(sf.train %>% as.data.frame %>% select(X,Y)))
    A_test <- inla.spde.make.A(mesh=mesh, loc=as.matrix(sf.test %>% as.data.frame %>% select(X,Y)))

    # Species occurrence
    y <- matrix(nrow=nrow(sf.train), ncol=length(variables) + 1)
    y[,i] <- sf.train[[species.take]]
    # Stack
    stk_y <- inla.stack(data=list(y=y),  
                            effects=list(c(list(ORG_STOF.z.field=1:spde$n.spde, 
                                                CN.z.field=1:spde$n.spde,
                                                N_totaal.z.field=1:spde$n.spde, 
                                                P_totaal.z.field=1:spde$n.spde, 
                                                K_CaCl2.z.field=1:spde$n.spde,
                                                u.field=1:spde$n.spde), list(intercept=1)),
                                         list(terrain=sf.train$terrain), 
                                         list(Dagnummer.z=sf.train$Dagnummer.z), 
                                         list(pH.z=sf.train$pH.z)),
                            A=list(A_train, 1, 1, 1),
                            tag="est.y")
    # Predicted over grid
    y <- matrix(nrow=nrow(sf.test), ncol=length(variables) + 1)
    stk_pred <- inla.stack(data=list(y=y),  
                            effects=list(c(list(ORG_STOF.z.field=1:spde$n.spde, 
                                                CN.z.field=1:spde$n.spde,
                                                N_totaal.z.field=1:spde$n.spde, 
                                                P_totaal.z.field=1:spde$n.spde, 
                                                K_CaCl2.z.field=1:spde$n.spde,
                                                u.field=1:spde$n.spde), list(intercept=1)),
                                         list(terrain=sf.test$terrain), 
                                         list(Dagnummer.z=sf.test$Dagnummer.z), 
                                         list(pH.z=sf.test$pH.z)),
                            A=list(A_test, 1, 1, 1),
                            tag="est.pred")

    stk <- inla.stack(stacks[["ORG_STOF.z"]], stacks[["CN.z"]], stacks[["N_totaal.z"]],
                      stacks[["P_totaal.z"]], stacks[["K_CaCl2.z"]], stk_y, stk_pred)

    formula <- y ~ -1 + intercept + ORG_STOF.z.intercept + CN.z.intercept + N_totaal.z.intercept + P_totaal.z.intercept + K_CaCl2.z.intercept +
      f(inla.group(terrain), model = "rw2") + f(inla.group(Dagnummer.z), model = "rw2") +  f(inla.group(pH.z), model = "rw2") + 
      f(ORG_STOF.z.i.field, model=spde) + f(ORG_STOF.z.field, copy="ORG_STOF.z.i.field", fixed=FALSE, hyper=list(theta=list(param=c(0, 0.001)))) +
      f(CN.z.i.field, model=spde) + f(CN.z.field, copy="CN.z.i.field", fixed=FALSE, hyper=list(theta=list(param=c(0, 0.001)))) +
      f(N_totaal.z.i.field, model=spde) + f(N_totaal.z.field, copy="N_totaal.z.i.field", fixed=FALSE, hyper=list(theta=list(param=c(0, 0.001))))  +
      f(P_totaal.z.i.field, model=spde) + f(P_totaal.z.field, copy="P_totaal.z.i.field", fixed=FALSE, hyper=list(theta=list(param=c(0, 0.001))))  +
      f(K_CaCl2.z.i.field, model=spde) + f(K_CaCl2.z.field, copy="K_CaCl2.z.i.field", fixed=FALSE, hyper=list(theta=list(param=c(0, 0.001)))) +
      f(u.field, model=spde)

    index.pred <- inla.stack.index(stk, tag="est.pred")$data

    families <- c(rep("gaussian", length(variables)), "binomial")
    output <-  tryCatch(
        {inla(formula, family=families, num.threads=1, #verbose=T,
              data=inla.stack.data(stk), 
              control.predictor=list(A=inla.stack.A(stk), compute=TRUE, link=i),
              control.inla = list(strategy = "adaptive", int.strategy = "eb"), control.compute=list(config = TRUE)) #int.strategy = "eb"
        }, error = function(cond) 
            {message(cond)
            return(NULL)}
    )

    # Second model results
    if (!is.null(output)) {
      # get inla posterior sample
      s1 <- inla.posterior.sample(400, output) 
      linear_predictor <- inla.posterior.sample.eval("APredictor", s1)[index.pred,]
      P = 1/(1+exp(-linear_predictor))
      sample_mean <- function(p) mean(rbinom(length(p), 1, p))
      area_prevalences <- apply(P, 2, sample_mean) #colMeans(1/(1+exp(-linear_predictor)))

      # Predicted vs observed
      predicted <- output$summary.fitted.values[index.pred,c("mean", "sd", "0.025quant", "0.975quant")]
      observed <- sf.test[[species.take]]

      predictions <- cbind(predicted, observed) %>% mutate(meta = "Uncertain", model="SDM", species=species.take, naam = provincie)
      prevalences <- data.frame(area_prevalences, meta = "Uncertain", model="SDM", species=species.take, naam = provincie)
      #write.table(predictions, "predictions_uncertain.csv", sep = ",", col.names = !file.exists("predictions_uncertain.csv"), row.names=F, append = T)
      #write.table(prevalences, "prevalences_uncertain.csv", sep = ",", col.names = !file.exists("prevalences_uncertain.csv"), row.names=F, append = T)
      fn.1 <- sprintf('predictions_uncertain/%s_%s.csv', species.take, provincie)
      fn.2 <- sprintf('prevalences_uncertain/%s_%s.csv', species.take, provincie)
      write.table(predictions, fn.1, sep = ",", col.names = T, row.names=F, append=F)
      write.table(prevalences, fn.2, sep = ",", col.names = T, row.names=F, append=F)
    }

  }
}
```

### FGR regions

Two-stage model (certain abiotics), save predictions and prevalences

```{r}
fgrs <- unique(sf.plots$Fys_geo_regio)
fgrs <- sort(fgrs[!(fgrs %in% c("gtw", "gtz"))]) # gtw (1) and gtz (5) samples
for (species.take in species) { 
  print(species.take)
  for (fgr in fgrs) {
    print(fgr)
    sf.train = sf.plots %>% filter(Fys_geo_regio != fgr) 
    sf.test = sf.plots %>% filter(Fys_geo_regio == fgr) 
    A_train <- inla.spde.make.A(mesh=mesh, loc=as.matrix(sf.train %>% as.data.frame %>% select(X,Y)))
    A_test <- inla.spde.make.A(mesh=mesh, loc=as.matrix(sf.test %>% as.data.frame %>% select(X,Y)))

    # Fit
    stk_y <- inla.stack(data = list(y=sf.train[[species.take]]),
                  A = list(A_train, 1, 1, 1, 1, 1, 1, 1, 1), 
                  effects = list(c(list(u.field=1:spde$n.spde), list(intercept=1)), 
                                 list(terrain=sf.train$terrain), 
                                 list(Dagnummer.z=sf.train$Dagnummer.z), 
                                 list(pH.z=sf.train$pH.z), 
                                 list(ORG_STOF.z=sf.train$ORG_STOF.z.mean),
                                 list(CN.z=sf.train$CN.z.mean), 
                                 list(N_totaal.z=sf.train$N_totaal.z.mean), 
                                 list(P_totaal.z=sf.train$P_totaal.z.mean), 
                                 list(K_CaCl2.z=sf.train$K_CaCl2.z.mean)
                                 ),
                  tag="est.y")

    stk_pred <- inla.stack(data=list(y=rep(NA, nrow(sf.test))),  
                            effects=list(c(list(u.field=1:spde$n.spde), list(intercept=1)),
                                         list(terrain=sf.test$terrain), 
                                         list(Dagnummer.z=sf.test$Dagnummer.z), 
                                         list(pH.z=sf.test$pH.z), 
                                         list(ORG_STOF.z=sf.test$ORG_STOF.z.mean),
                                         list(CN.z=sf.test$CN.z.mean), 
                                         list(N_totaal.z=sf.test$N_totaal.z.mean), 
                                         list(P_totaal.z=sf.test$P_totaal.z.mean), 
                                         list(K_CaCl2.z=sf.test$K_CaCl2.z.mean)
                                 ),
                            A=list(A_test, 1, 1, 1, 1, 1, 1, 1, 1),
                            tag="est.pred")

    stk <- inla.stack(stk_y, stk_pred)

    formula <- y ~ -1 + intercept + f(inla.group(terrain), model = "rw2") + 
      f(inla.group(Dagnummer.z), model = "rw2") +  f(inla.group(pH.z), model = "rw2") +
      CN.z + ORG_STOF.z + N_totaal.z + P_totaal.z + K_CaCl2.z + f(u.field, model=spde)

    index.pred <- inla.stack.index(stk, tag="est.pred")$data

    output <-  tryCatch(
        {inla(formula, num.threads=1,
                    data=inla.stack.data(stk, spde=spde), family="binomial",
                    control.predictor=list(A=inla.stack.A(stk), compute=TRUE, link=1), 
                    control.inla = list(strategy = "adaptive", int.strategy = "eb"), control.compute=list(config = TRUE)) #int.strategy = "eb"
        }, error = function(cond) 
            {message(cond)
            return(NULL)}
    )

    # First model results
    if (!is.null(output)) {
      # get inla posterior sample
      s1 <- inla.posterior.sample(400, output) 
      linear_predictor <- inla.posterior.sample.eval("APredictor", s1)[index.pred,]
      P = 1/(1+exp(-linear_predictor))
      sample_mean <- function(p) mean(rbinom(length(p), 1, p))
      area_prevalences <- apply(P, 2, sample_mean) #colMeans(1/(1+exp(-linear_predictor)))

      # Predicted vs observed
      predicted <- output$summary.fitted.values[index.pred,c("mean", "sd", "0.025quant", "0.975quant")]
      observed <- sf.test[[species.take]]

      predictions <- cbind(predicted, observed) %>% mutate(meta = "Certain", model="SDM", species=species.take, naam = fgr)
      prevalences <- data.frame(area_prevalences, meta = "Certain", model="SDM", species=species.take, naam = fgr)
      #write.table(predictions, "predictions_certain2.csv", sep = ",", col.names = !file.exists("predictions_certain2.csv"), row.names=F, append = T)
      #write.table(prevalences, "prevalences_certain2.csv", sep = ",", col.names = !file.exists("prevalences_certain2.csv"), row.names=F, append = T)
      fn.1 <- sprintf('predictions_certain2/%s_%s.csv', species.take, fgr)
      fn.2 <- sprintf('prevalences_certain2/%s_%s.csv', species.take, fgr)
      write.table(predictions, fn.1, sep = ",", col.names = T, row.names=F, append=F)
      write.table(prevalences, fn.2, sep = ",", col.names = T, row.names=F, append=F)
    }

  }
}
```

Joint model (uncertain abiotics), save predictions and prevalences

```{r}
fgrs <- unique(sf.plots$Fys_geo_regio)
fgrs <- sort(fgrs[!(fgrs %in% c("gtw", "gtz"))]) # gtw (1) and gtz (5) samples
for (species.take in species) { 
  print(species.take)
  for (fgr in fgrs) {
    print(fgr)
    sf.train = sf.plots %>% filter(Fys_geo_regio != fgr) 
    sf.test = sf.plots %>% filter(Fys_geo_regio == fgr) 
    A_train <- inla.spde.make.A(mesh=mesh, loc=as.matrix(sf.train %>% as.data.frame %>% select(X,Y)))
    A_test <- inla.spde.make.A(mesh=mesh, loc=as.matrix(sf.test %>% as.data.frame %>% select(X,Y)))

    # Species occurrence
    y <- matrix(nrow=nrow(sf.train), ncol=length(variables) + 1)
    y[,i] <- sf.train[[species.take]]
    # Stack
    stk_y <- inla.stack(data=list(y=y),  
                            effects=list(c(list(ORG_STOF.z.field=1:spde$n.spde, 
                                                CN.z.field=1:spde$n.spde,
                                                N_totaal.z.field=1:spde$n.spde, 
                                                P_totaal.z.field=1:spde$n.spde, 
                                                K_CaCl2.z.field=1:spde$n.spde,
                                                u.field=1:spde$n.spde), list(intercept=1)),
                                         list(terrain=sf.train$terrain), 
                                         list(Dagnummer.z=sf.train$Dagnummer.z), 
                                         list(pH.z=sf.train$pH.z)),
                            A=list(A_train, 1, 1, 1),
                            tag="est.y")
    # Predicted over grid
    y <- matrix(nrow=nrow(sf.test), ncol=length(variables) + 1)
    stk_pred <- inla.stack(data=list(y=y),  
                            effects=list(c(list(ORG_STOF.z.field=1:spde$n.spde, 
                                                CN.z.field=1:spde$n.spde,
                                                N_totaal.z.field=1:spde$n.spde, 
                                                P_totaal.z.field=1:spde$n.spde, 
                                                K_CaCl2.z.field=1:spde$n.spde,
                                                u.field=1:spde$n.spde), list(intercept=1)),
                                         list(terrain=sf.test$terrain), 
                                         list(Dagnummer.z=sf.test$Dagnummer.z), 
                                         list(pH.z=sf.test$pH.z)),
                            A=list(A_test, 1, 1, 1),
                            tag="est.pred")

    stk <- inla.stack(stacks[["ORG_STOF.z"]], stacks[["CN.z"]], stacks[["N_totaal.z"]],
                      stacks[["P_totaal.z"]], stacks[["K_CaCl2.z"]], stk_y, stk_pred)

    formula <- y ~ -1 + intercept + ORG_STOF.z.intercept + CN.z.intercept + N_totaal.z.intercept + P_totaal.z.intercept + K_CaCl2.z.intercept +
      f(inla.group(terrain), model = "rw2") + f(inla.group(Dagnummer.z), model = "rw2") +  f(inla.group(pH.z), model = "rw2") + 
      f(ORG_STOF.z.i.field, model=spde) + f(ORG_STOF.z.field, copy="ORG_STOF.z.i.field", fixed=FALSE, hyper=list(theta=list(param=c(0, 0.001)))) +
      f(CN.z.i.field, model=spde) + f(CN.z.field, copy="CN.z.i.field", fixed=FALSE, hyper=list(theta=list(param=c(0, 0.001)))) +
      f(N_totaal.z.i.field, model=spde) + f(N_totaal.z.field, copy="N_totaal.z.i.field", fixed=FALSE, hyper=list(theta=list(param=c(0, 0.001))))  +
      f(P_totaal.z.i.field, model=spde) + f(P_totaal.z.field, copy="P_totaal.z.i.field", fixed=FALSE, hyper=list(theta=list(param=c(0, 0.001))))  +
      f(K_CaCl2.z.i.field, model=spde) + f(K_CaCl2.z.field, copy="K_CaCl2.z.i.field", fixed=FALSE, hyper=list(theta=list(param=c(0, 0.001)))) +
      f(u.field, model=spde)

    index.pred <- inla.stack.index(stk, tag="est.pred")$data

    families <- c(rep("gaussian", length(variables)), "binomial")
    output <-  tryCatch(
        {inla(formula, family=families, num.threads=1, #verbose=T,
              data=inla.stack.data(stk), 
              control.predictor=list(A=inla.stack.A(stk), compute=TRUE, link=i),
              control.inla = list(strategy = "adaptive", int.strategy = "eb"), control.compute=list(config = TRUE)) #int.strategy = "eb"
        }, error = function(cond) 
            {message(cond)
            return(NULL)}
    )

    # Second model results
    if (!is.null(output)) {
      # get inla posterior sample
      s1 <- inla.posterior.sample(400, output) 
      linear_predictor <- inla.posterior.sample.eval("APredictor", s1)[index.pred,]
      P = 1/(1+exp(-linear_predictor))
      sample_mean <- function(p) mean(rbinom(length(p), 1, p))
      area_prevalences <- apply(P, 2, sample_mean) #colMeans(1/(1+exp(-linear_predictor)))

      # Predicted vs observed
      predicted <- output$summary.fitted.values[index.pred,c("mean", "sd", "0.025quant", "0.975quant")]
      observed <- sf.test[[species.take]]

      predictions <- cbind(predicted, observed) %>% mutate(meta = "Uncertain", model="SDM", species=species.take, naam = fgr)
      prevalences <- data.frame(area_prevalences, meta = "Uncertain", model="SDM", species=species.take, naam = fgr)
      #write.table(predictions, "predictions_uncertain2.csv", sep = ",", col.names = !file.exists("predictions_uncertain2.csv"), row.names=F, append = T)
      #write.table(prevalences, "prevalences_uncertain2.csv", sep = ",", col.names = !file.exists("prevalences_uncertain2.csv"), row.names=F, append = T)
      fn.1 <- sprintf('predictions_uncertain2/%s_%s.csv', species.take, fgr)
      fn.2 <- sprintf('prevalences_uncertain2/%s_%s.csv', species.take, fgr)
      write.table(predictions, fn.1, sep = ",", col.names = T, row.names=F, append=F)
      write.table(prevalences, fn.2, sep = ",", col.names = T, row.names=F, append=F)
    }

  }
}
```

## Combine and export data

Combine and export the results of fitting all models to data:

```{r}
# Model parameters and predictions in entire data set
c(list.files("results_certain", full.names=T),
  list.files("results_uncertain", full.names=T)) %>% 
  map_dfr(read_csv, show_col_types=F) %>% 
  mutate(species=replace_na(species, ""))  %>% 
  write.csv(file="data_paper/results.csv", row.names=F)
# Predicted occurrence and prevalence in validation data set (Provinces)
c(list.files("predictions_certain", full.names=T), 
  list.files("predictions_uncertain", full.names=T)) %>% 
  map_dfr(read_csv, show_col_types=F) %>% 
  mutate(species=replace_na(species, ""))  %>% 
  write.csv(file="data_paper/predictions_province.csv", row.names=F)
c(list.files("prevalences_certain", full.names=T), 
  list.files("prevalences_uncertain", full.names=T)) %>% 
  map_dfr(read_csv, show_col_types=F) %>% 
  mutate(species=replace_na(species, ""))  %>% 
  write.csv(file="data_paper/prevalences_province.csv", row.names=F)
# Predicted occurrence and prevalence in validation data set (FGR regions)
c(list.files("predictions_certain2", full.names=T), 
  list.files("predictions_uncertain2", full.names=T)) %>% 
  map_dfr(read_csv, show_col_types=F) %>% 
  mutate(species=replace_na(species, ""))  %>% 
  write.csv(file="data_paper/predictions_fgr.csv", row.names=F)
c(list.files("prevalences_certain2", full.names=T), 
  list.files("prevalences_uncertain2", full.names=T)) %>% 
  map_dfr(read_csv, show_col_types=F) %>% 
  mutate(species=replace_na(species, ""))  %>% 
  write.csv(file="data_paper/prevalences_fgr.csv", row.names=F)
```

## Results: crowberry example

```{r}
results <- read.csv("data_paper/results.csv")
#results <- read.csv("data_paper/results.csv", col.names=c("species", "meta", "model", "term", "ID", "mean", "sd", "X0.025quant", "X0.5quant", "X0.975quant")) # LSF cluster results
```

```{r}
species.take <- "Kraaihei" #Gewoon reukgras, Engels raaigras, Pitrus, Duizendbland, Jacobskruiskruid
terms <- results %>% filter((species == species.take) | (species == "")) #is.na(species)
```

Spline parameters (SDM)

```{r}
# Day number, pH, terrain
terms.plot <- terms %>% filter(term %in% c("pH.z", "Dagnummer.z", "terrain")) %>% mutate(ID=ID %>% as.numeric) %>%
  mutate(term=recode(term, "pH.z" = "pH (z-score)", "Dagnummer.z"="Day number (months)", "terrain"="Terrain (more - less natural)")) %>%
  mutate(meta=factor(recode(meta, "Certain" = "Two-Stage", "Uncertain"="Joint"), levels=c("Two-Stage", "Joint")))
plot.splines <- ggplot(terms.plot, aes(x=ID, y=mean, col=meta, group=meta)) + 
  geom_point(position=position_dodge(width=0.5), size=2) + geom_line() +
  geom_errorbar(position=position_dodge(width=0.5), aes(ymin=X0.025quant, ymax=X0.975quant)) +
  theme(legend.position="top", axis.text.x = element_text(angle = 90)) + 
  scale_color_discrete(name = "Model") +
  scale_x_continuous(breaks=seq(-4,4), name='Covariate Value') + 
  scale_y_continuous(name='Value') +
  facet_wrap( ~ term, ncol = 10, scales="free")
plot.splines
```

Abiotic parameters (SDM)

```{r}
# Linear terms
terms.plot <- terms %>% filter((term == "fixed") & (ID != "intercept")) %>% 
  mutate(ID=fct_relevel(ID, "ORG_STOF.z", "CN.z", "N_totaal.z", "P_totaal.z", "K_CaCl2.z")) %>% #"intercept", 
  mutate(term=recode(term, "fixed" = "Abiotic (z-score)")) %>%
  mutate(meta=factor(recode(meta, "Certain" = "Two-Stage", "Uncertain"="Joint"), levels=c("Two-Stage", "Joint")))
plot.abiotic <- ggplot(terms.plot, aes(x=ID, y=mean, col=meta, group=meta)) + 
  geom_point(position=position_dodge(width=0.5), size=2) +
  geom_errorbar(position=position_dodge(width=0.5), aes(ymin=X0.025quant, ymax=X0.975quant)) +
  theme(legend.position="top", axis.text.x = element_text(angle = 90)) + 
  scale_color_discrete(name = "Model Abiotics") +
  scale_x_discrete(name='Covariate') + 
  scale_y_continuous(name='Value') + #coord_cartesian(ylim=c(-5,5))+ #ylim=c(-10,10)
  facet_wrap( ~ term, ncol = 10, scales="free")
plot.abiotic
```

Spatial component parameters (SDM)

```{r}
# Spatial terms
terms.plot <- terms %>% filter((model == "SDM") & (term == "spatial")) %>% 
  mutate(ID=fct_relevel(ID, "log.range", "log.variance")) %>%
  mutate(term=recode(term, "spatial" = "Spatial (km)")) %>%
  mutate(meta=factor(recode(meta, "Certain" = "Two-Stage", "Uncertain"="Joint"), levels=c("Two-Stage", "Joint")))
plot.spatial <- ggplot(terms.plot, aes(x=ID, y=mean, col=meta, group=meta)) + 
  geom_point(position=position_dodge(width=0.5), size=2) +
  geom_errorbar(position=position_dodge(width=0.5), aes(ymin=X0.025quant, ymax=X0.975quant)) +
  theme(legend.position="top", axis.text.x = element_text(angle = 90)) + 
  scale_color_discrete(name = "Model Abiotics") +
  scale_x_discrete(name='Covariate') + 
  scale_y_continuous(name='Value') + #+coord_cartesian(ylim=c(-5,5))+ #ylim=c(-10,10)
  facet_wrap( ~ term, ncol = 10, scales="free")
plot.spatial
```

Spatial component parameters (abiotics)

```{r}
# Spatial terms (abiotics)
terms.plot <- terms %>% filter((term =="spatial") & (model != "SDM")) %>%
  mutate(ID=fct_relevel(ID, "log.range", "log.variance", "log.variance.eps")) %>%
  mutate(term=recode(term, "spatial" = "Spatial (km)"))%>%
  mutate(meta=factor(recode(meta, "Certain" = "Two-Stage", "Uncertain"="Joint"), levels=c("Two-Stage", "Joint")))
plot.extrapolate <- ggplot(terms.plot, aes(x=ID, y=mean, col=meta, group=meta)) + 
  geom_point(position=position_dodge(width=0.5), size=2) +
  geom_errorbar(position=position_dodge(width=0.5), aes(ymin=X0.025quant, ymax=X0.975quant)) +
  theme(legend.position="top", axis.text.x = element_text(angle = 90)) + 
  scale_color_discrete(name = "Model Abiotics") +
  scale_x_discrete(name='Covariate') + 
  scale_y_continuous(name='Value') +
  facet_wrap( ~ model, ncol = 10, scales="free")
plot.extrapolate
```

Combine all parameters to a single plot

```{r}
plot.all <- ggdraw() +
  draw_plot(plot.splines, x = 0, y = .666, width = 1, height = .333) +
  draw_plot(plot.abiotic + theme(legend.position = "none"), x = 0, y = .333, width = .77, height = .333) +
  draw_plot(plot.spatial + theme(legend.position = "none"), x = 0.77, y = .333, width = 0.23, height = 0.333) +
  draw_plot(plot.extrapolate + theme(legend.position = "none"), x = 0, y = 0, width = 1, height = 0.333) +
  draw_plot_label(label = c("A", "B", "C", "D"), size = 15, x = c(0.03, 0.03, 0.79, 0.03), y = c(0.96, 0.70, 0.70, 0.37))
plot.all
ggsave('results_paper/species_parameters.png', width=2000, heigh=2500, units='px')
```

Map of observed vs. predicted SDM from both models, probability scale

```{r}
# Predictions
terms.plot <- terms %>% filter((model == "SDM") & (term == "Predict")) %>% mutate(Plot=ID %>% as.factor)

# Combine data observations with model predictions
sf.base <- sf.grid %>% select(Plot, geometry.1) %>% st_set_geometry("geometry.1")
terms.y <- sf.base %>% st_join(sf.plots %>% select(all_of(species.take)), st_contains) %>% 
  group_by(Plot) %>% summarize(y=mean(get(species.take), na.rm=T)) %>% as.data.frame %>% select(Plot, y)
terms.sf <- sf.base %>% merge(terms.plot) %>% merge(terms.y)

plot0 <- ggplot(NL) + geom_sf() + geom_sf(aes(fill=y), data=terms.sf %>% filter(meta=="Uncertain")) + 
  scale_fill_viridis_c(option="B", limits=c(0,1)) + ggtitle("Raw data") + 
  theme(legend.position="bottom", legend.key.width = unit(0.06, "npc"))
plot1 <- ggplot(NL) + geom_sf() + geom_sf(aes(fill=mean), data=terms.sf %>% filter(meta=="Uncertain")) + 
  scale_fill_viridis_c(option="B", limits=c(0,1)) + ggtitle("Joint Model") + 
  theme(legend.position="bottom", legend.key.width = unit(0.06, "npc"))
plot2 <- ggplot(NL) + geom_sf() + geom_sf(aes(fill=mean), data=terms.sf %>% filter(meta=="Certain")) + 
  scale_fill_viridis_c(option="B", limits=c(0,1)) + ggtitle("Two-Stage Model") + 
  theme(legend.position="bottom", legend.key.width = unit(0.06, "npc"))
plot4 <- ggplot(NL) + geom_sf() + geom_sf(aes(fill=sd), data=terms.sf %>% filter(meta=="Uncertain")) + 
  scale_fill_gradient(low='darkblue', high='yellow', limits=c(0,max(terms.sf$sd))) + ggtitle("Joint Model") + 
  theme(legend.position="bottom", legend.key.width = unit(0.06, "npc"))
plot5 <- ggplot(NL) + geom_sf() + geom_sf(aes(fill=sd), data=terms.sf %>% filter(meta=="Certain")) + 
  scale_fill_gradient(low='darkblue', high='yellow', limits=c(0,max(terms.sf$sd))) + ggtitle("Two-Stage Model") + 
  theme(legend.position="bottom", legend.key.width = unit(0.06, "npc"))
#plot.all <- plot_grid(plot0, plot1, plot2, NULL, plot4, plot5, nrow = 2) + theme(plot.background = element_rect(fill = "white", colour = NA))
plot.all <- ggdraw() +
  draw_plot(plot0, x = 0, y = .5, width = 0.33, height = .5) +
  draw_plot(plot1, x = 0.33, y = .5, width = 0.33, height = .5) +
  draw_plot(plot2, x = 0.66, y = .5, width = 0.33, height = .5) +
  draw_plot(plot4, x = 0.33, y = 0, width = 0.33, height = .5) +
  draw_plot(plot5, x = 0.66, y = 0, width = 0.33, height = .5) +
  draw_plot_label(label = c("A", "B", "C", "D", "E"), size = 15, x = c(0.00, 0.33, 0.66, 0.33, 0.66), y = c(1.00, 1.00, 1.00, 0.50, 0.50))
plot.all
ggsave('results_paper/species_SDM_y.png', width=3200, heigh=3000, units='px')
```

Map of observed vs. predicted SDM from both models, log-odds scale

```{r}
# Predictions (log-odds)
terms.plot <- terms %>% filter((model == "SDM") & (term == "APredict")) %>% mutate(Plot=ID %>% as.factor)
y.mean <- mean(sf.plots[[species.take]], na.rm=T)
y.mean <- log(y.mean/(1-y.mean))

# Combine data observations with model predictions
sf.base <- sf.grid %>% select(Plot, geometry.1) %>% st_set_geometry("geometry.1")
terms.y <- sf.base %>% st_join(sf.plots %>% select(all_of(species.take)), st_contains) %>% 
  group_by(Plot) %>% summarize(y=mean(get(species.take), na.rm=T)) %>% as.data.frame %>% select(Plot, y)
terms.sf <- sf.base %>% merge(terms.plot) %>% merge(terms.y)

plot0 <- ggplot(NL) + geom_sf() + geom_sf(aes(fill=y), data=terms.sf %>% filter(meta=="Uncertain")) + 
  scale_fill_viridis_c(option="B", limits=c(0,1)) + ggtitle("Raw data") + 
  theme(legend.position="bottom", legend.key.width = unit(0.06, "npc"))
plot1 <- ggplot(NL) + geom_sf() + geom_sf(aes(fill=mean), data=terms.sf %>% filter(meta=="Uncertain")) + 
  scale_fill_continuous_divergingx(palette = 'PiYG', mid = y.mean) + ggtitle("Joint Model") + 
  theme(legend.position="bottom", legend.key.width = unit(0.06, "npc"))
plot2 <- ggplot(NL) + geom_sf() + geom_sf(aes(fill=mean), data=terms.sf %>% filter(meta=="Certain")) + 
  scale_fill_continuous_divergingx(palette = 'PiYG', mid = y.mean) + ggtitle("Two-Stage Model") + 
  theme(legend.position="bottom", legend.key.width = unit(0.06, "npc"))
plot4 <- ggplot(NL) + geom_sf() + geom_sf(aes(fill=sd), data=terms.sf %>% filter(meta=="Uncertain")) + 
  scale_fill_gradient(low='darkblue', high='orange', limits=c(0,max(terms.sf$sd))) + ggtitle("Joint Model") + 
  theme(legend.position="bottom", legend.key.width = unit(0.06, "npc"))
plot5 <- ggplot(NL) + geom_sf() + geom_sf(aes(fill=sd), data=terms.sf %>% filter(meta=="Certain")) + 
  scale_fill_gradient(low='darkblue', high='orange', limits=c(0,max(terms.sf$sd))) + ggtitle("Two-Stage Model") + 
  theme(legend.position="bottom", legend.key.width = unit(0.06, "npc"))
#plot.all <- plot_grid(plot0, plot1, plot2, NULL, plot4, plot5, nrow = 2) + theme(plot.background = element_rect(fill = "white", colour = NA))
plot.all <- ggdraw() +
  draw_plot(plot0, x = 0, y = .5, width = 0.33, height = .5) +
  draw_plot(plot1, x = 0.33, y = .5, width = 0.33, height = .5) +
  draw_plot(plot2, x = 0.66, y = .5, width = 0.33, height = .5) +
  draw_plot(plot4, x = 0.33, y = 0, width = 0.33, height = .5) +
  draw_plot(plot5, x = 0.66, y = 0, width = 0.33, height = .5) +
  draw_plot_label(label = c("A", "B", "C", "D", "E"), size = 15, x = c(0.00, 0.33, 0.66, 0.33, 0.66), y = c(1.00, 1.00, 1.00, 0.50, 0.50))
plot.all
ggsave('results_paper/species_SDM_logodds.png', width=3200, heigh=3000, units='px')

```

Map of observed vs. predicted abiotics from both models

```{r}
for (var in c("ORG_STOF.z", "CN.z", "N_totaal.z", "P_totaal.z", "K_CaCl2.z")) {#var <- "ORG_STOF.z"
# Predictions
terms.plot <- terms %>% filter((model == var) & (term == "Predict")) %>% mutate(Plot=ID %>% as.factor)

# Combine data observations with model predictions
sf.base <- sf.grid %>% select(Plot, geometry.1) %>% st_set_geometry("geometry.1")
terms.y <- sf.base %>% st_join(sf.plots[var], st_contains) %>% 
  group_by(Plot) %>% summarize(y=mean(get(var), na.rm=T)) %>% as.data.frame %>% select(Plot, y)
terms.sf <- sf.base %>% merge(terms.plot) %>% merge(terms.y)

plot0 <- ggplot(NL) + geom_sf() + geom_sf(aes(fill=y), data=terms.sf %>% filter(meta=="Uncertain")) + 
  scale_fill_gradient2(low='blue', midpoint=0, high='red', limits=c(-3,3)) + 
  ggtitle(sprintf('Raw data %s', var)) + theme(legend.position="bottom", legend.key.width = unit(0.06, "npc"))
plot1 <- ggplot(NL) + geom_sf() + geom_sf(aes(fill=mean), data=terms.sf %>% filter(meta=="Uncertain")) + 
  scale_fill_gradient2(low='blue', midpoint=0, high='red', limits=c(-3,3)) + 
  ggtitle(sprintf('Joint Model %s', var)) + theme(legend.position="bottom", legend.key.width = unit(0.06, "npc"))
plot2 <- ggplot(NL) + geom_sf() + geom_sf(aes(fill=mean), data=terms.sf %>% filter(meta=="Certain")) + 
  scale_fill_gradient2(low='blue', midpoint=0, high='red', limits=c(-3,3)) + 
  ggtitle(sprintf('Two-Stage Model %s', var)) + theme(legend.position="bottom", legend.key.width = unit(0.06, "npc"))
plot4 <- ggplot(NL) + geom_sf() + geom_sf(aes(fill=sd), data=terms.sf %>% filter(meta=="Uncertain")) + 
  scale_fill_gradient(low='white', high='red', limits=c(0,max(terms.sf$sd))) + 
  ggtitle(sprintf('Joint Model %s', var)) + theme(legend.position="bottom", legend.key.width = unit(0.06, "npc"))
plot5 <- ggplot(NL) + geom_sf() + geom_sf(aes(fill=sd), data=terms.sf %>% filter(meta=="Certain")) + 
  scale_fill_gradient(low='white', high='red', limits=c(0,max(terms.sf$sd))) + 
  ggtitle(sprintf('Two-Stage Model %s', var)) + theme(legend.position="bottom", legend.key.width = unit(0.06, "npc"))
#plot.all <- plot_grid(plot0, plot1, plot2, NULL, plot4, plot5, nrow = 2) + theme(plot.background = element_rect(fill = "white", colour = NA))
plot.all <- ggdraw() +
  draw_plot(plot0, x = 0, y = .5, width = 0.33, height = .5) +
  draw_plot(plot1, x = 0.33, y = .5, width = 0.33, height = .5) +
  draw_plot(plot2, x = 0.66, y = .5, width = 0.33, height = .5) +
  draw_plot(plot4, x = 0.33, y = 0, width = 0.33, height = .5) +
  draw_plot(plot5, x = 0.66, y = 0, width = 0.33, height = .5) +
  draw_plot_label(label = c("A", "B", "C", "D", "E"), size = 15, x = c(0.00, 0.33, 0.66, 0.33, 0.66), y = c(1.00, 1.00, 1.00, 0.50, 0.50))
ggsave(sprintf('results_paper/species_%s.png', var), width=3200, heigh=3000, units='px')
}
```

Scatter plot of predicted values in each grid cell: two-stage vs joint model

```{r}
terms.plot.1 <- terms %>% filter((model == "SDM") & (term == "APredict")) %>% mutate(Plot=ID %>% as.factor) %>% 
  mutate(meta=factor(recode(meta, "Certain" = "Two-Stage Model", "Uncertain"="Joint Model"), levels=c("Two-Stage Model", "Joint Model"))) %>%
  select(species,meta,model,term,ID,mean,sd) %>% pivot_longer(c(mean,sd), names_to="type", values_to="estimate") %>% 
  pivot_wider(names_from = meta, values_from = estimate)

terms.plot.2 <- terms %>% filter((model != "SDM") & (term == "Predict")) %>% 
  mutate(meta=factor(recode(meta, "Certain" = "Two-Stage Model", "Uncertain"="Joint Model"), levels=c("Two-Stage Model", "Joint Model"))) %>%
  select(meta,model,term,ID,mean,sd) %>% pivot_longer(c(mean,sd), names_to="type", values_to="estimate") %>% 
  pivot_wider(names_from = meta, values_from = estimate) %>% mutate(species=species.take)

terms.plot <- rbind(terms.plot.1, terms.plot.2) %>% 
  mutate(model=fct_relevel(model, "SDM", "ORG_STOF.z", "CN.z", "N_totaal.z", "P_totaal.z", "K_CaCl2.z"))

plot.scatter <- ggplot() + geom_point(aes(x=`Two-Stage Model`, y=`Joint Model`), data=terms.plot, size=0.1, alpha=0.2) + geom_abline() + 
     facet_wrap( ~ type + model, ncol = 3, scales="free") #+ ggtitle('Predicted Log-Odds (SDM) and \nPredicted Mean (Abiotics) at Grid Cells')
plot.scatter
ggsave('results_paper/species_scatter.png', width=1800, heigh=2400, units='px')
```

## Results: all species

Abiotic parameters for all species two-stage vs. joint model

```{r}
# Linear terms for all species
terms.plot <- results %>% filter((term == "fixed") & (model == "SDM")  & (ID != "intercept")) %>% 
  mutate(ID=fct_relevel(ID, "ORG_STOF.z", "CN.z", "N_totaal.z", "P_totaal.z", "K_CaCl2.z")) %>% #"intercept",  
  mutate(meta=factor(recode(meta, "Certain" = "Two-Stage", "Uncertain"="Joint"), levels=c("Two-Stage", "Joint"))) %>%
  merge(species.map) %>% mutate(species = name)
plot.abiotics.all <- ggplot(terms.plot, aes(x=ID, y=mean, col=meta, group=meta)) + 
  geom_point(position=position_dodge(width=0.5), size=2) +
  geom_errorbar(position=position_dodge(width=0.5), aes(ymin=X0.025quant, ymax=X0.975quant)) +
  theme(legend.position="top", axis.text.x = element_text(angle = 90)) + 
  scale_color_discrete(name = "Model") +
  scale_x_discrete(name='Covariate') + 
  scale_y_continuous(name='Value') +coord_cartesian(ylim=c(-5,5))+ #ylim=c(-10,10)
  facet_wrap( ~ species, ncol = 5)#, scales="free"
plot.abiotics.all
ggsave('results_paper/all_abiotics.png', width=2400, heigh=4000, units='px')
```

Table of correlations between predictions in each grid cell of two-stage vs. joint model

```{r}
# Uncertain vs Certain abiotic predictions
certain.abiotics <- results %>% filter(species == "") %>% transmute(model,ID,Certain=mean)
uncertain.abiotics <- results %>% filter((meta == "Uncertain") & (model != "SDM") & (term == "Predict")) %>% transmute(species, model,ID,Uncertain=mean)
abiotics <- uncertain.abiotics %>% merge(certain.abiotics)
# Uncertain vs Certain SDM predictions
certain.sdm <- results %>% filter((meta == "Certain") & (model == "SDM") & (term == "APredict")) %>% transmute(species, model,ID,Certain=mean)
uncertain.sdm <- results %>% filter((meta == "Uncertain") & (model == "SDM") & (term == "APredict")) %>% transmute(species, model,ID,Uncertain=mean)
sdm <- certain.sdm %>% merge(uncertain.sdm)
# Table of correlations
compare <- rbind(abiotics, sdm)
compare.cor <- compare %>% group_by(species, model) %>% 
  summarize(cor = sprintf("%.2f", round(cor(Certain, Uncertain),2))) %>% 
  pivot_wider(names_from=model, values_from=cor) %>%
  merge(species.map) %>% mutate(species = name) %>% select(-name) %>% arrange(species)
compare.cor %>% write.table('results_paper/correlation.csv', sep=',', row.names=F)
```

Histogram of these correlations

```{r}
compare.cor <- compare %>% group_by(species, model) %>% 
  summarize(cor = cor(Certain, Uncertain)) %>%
  mutate(model=fct_relevel(model, "SDM", "ORG_STOF.z", "CN.z", "N_totaal.z", "P_totaal.z", "K_CaCl2.z"))

ggplot(compare.cor, aes(x=cor)) + 
  geom_histogram(color="black", fill="white", breaks=seq(0.8,1.0,0.01)) + 
  ylab("Number of Species") +xlab("Pearson Correlation Coefficient") +
  facet_wrap( ~ model, ncol = 3)  #+ggtitle('Correlation between Certain vs. Uncertain Abiotic Model Predictions')
ggsave('results_paper/species_correlation.png', width=2000, heigh=1200, units='px')

```

Table of correlations between abiotic parameters in each species of two-stage vs. joint model

```{r}
# Uncertain vs Certain abiotic predictions
certain.abiotics <- results %>% filter((meta == "Certain") & (model == "SDM") & (term == "fixed") & (ID != "intercept")) %>%
  transmute(species, model,ID,Certain=mean)
uncertain.abiotics <- results %>% filter((meta == "Uncertain") & (model == "SDM") & (term == "fixed") & (ID != "intercept")) %>%
  transmute(species, model,ID,Uncertain=mean)
abiotics <- uncertain.abiotics %>% merge(certain.abiotics)
# Table of correlations
compare.cor <- abiotics %>% group_by(ID) %>% summarize(cor = round(cor(Certain, Uncertain),2)) %>% 
  pivot_wider(names_from=ID, values_from=cor)
compare.cor
```

Validation set illustration

```{r}
# Leave province out
sf.grid$provincie <- as.factor(sf.grid$naam)
plot.validation.1 <- ggplot(sf.grid) +
	geom_sf(aes(geometry=geometry.1, fill=provincie)) + 
	geom_sf(fill=NA, data=NL) +
  scale_fill_discrete(name = "Province") +
	ggtitle('Leave-province-out')

# Leave FGR region out
sf.grid$fgr <- factor(sf.grid$lmf_indeling, levels=c("afz", "duo", "duw", "hll", "hzn", "hzo", "hzv",
                                                     "hzz", "lvh", "lvn", "riv", "zkm", "zkn", "zkz"))
plot.validation.2 <- ggplot(sf.grid) +
	geom_sf(aes(geometry=geometry.1, fill=fgr)) + 
	geom_sf(fill=NA, data=NL) +
  scale_fill_manual(name = "FGR", values=c("afz"="lightseagreen", "duo"="darkgoldenrod", "duw"="darkgoldenrod1", "hll"="forestgreen", 
                                           "hzn"="wheat1", "hzo"="wheat2", "hzv"="wheat3", "hzz"="wheat4", 
                                           "lvh"="darkseagreen", "lvn"="darkseagreen1", "riv"="steelblue", 
                                           "zkm"="snow2", "zkn"="snow3", "zkz"="snow4")) +
	ggtitle('Leave-FGR-out')

# Combine plots
plot.all <- ggdraw() +
  draw_plot(plot.validation.1, x = 0, y = 0, width = 0.53, height = 1) +
  draw_plot(plot.validation.2, x = 0.53, y = 0, width = 0.47, height = 1) +
  draw_plot_label(label = c("A", "B"), size = 15, x = c(0.01, 0.53), y = c(0.99,0.99)) +
  theme(plot.background = element_rect(fill = "white", colour = NA))
plot.all
ggsave('results_paper/leave_out_validation.png', width=2500, heigh=1300, units='px')

```

## Results: training and validation

Calculate overall RMSE in each province

```{r}
predictions <- read.csv("data_paper/predictions_province.csv")
#predictions <- read.csv("data_paper/predictions_province.csv", col.names=c("mean", "sd", "X0.025quant", "X0.975quant", "observed", "meta", "model", "species", "naam")) # LSF cluster results
rmse.1 <- predictions %>% group_by(meta, species) %>% summarize(rmse=sqrt(mean((observed-mean)**2, na.rm=T))) 
```

Calculate overall RMSE in each FGR region

```{r}
predictions <- read.csv("data_paper/predictions_fgr.csv")
#predictions <- read.csv("data_paper/predictions_fgr.csv", col.names=c("mean", "sd", "X0.025quant", "X0.975quant", "observed", "meta", "model", "species", "naam")) # LSF cluster results
rmse.2 <- predictions %>% group_by(meta, species) %>% summarize(rmse=sqrt(mean((observed-mean)**2, na.rm=T))) 
```

Scatter plot of two-stage vs. joint model overall RMSE in each species:

```{r}
rmse.both <- rbind(rmse.1 %>% mutate(validation="Province"),
                   rmse.2 %>% mutate(validation="FGR")) %>%
  mutate(meta=factor(recode(meta, "Certain" = "Two-Stage Model", "Uncertain"="Joint Model"), levels=c("Two-Stage Model", "Joint Model")))
rmse.both %>% pivot_wider(names_from = c(validation, meta), values_from = rmse) %>% 
  mutate(across(2:5, function(x) sprintf("%.3f", round(x,3)))) %>% 
  merge(species.map) %>% mutate(species = name) %>% select(-name) %>% arrange(species)%>%
  write.table('results_paper/rmse_compare.csv', sep=',', row.names=F)

ggplot(rmse.both  %>% pivot_wider(names_from = meta, values_from = rmse)) + 
  geom_point(aes(x=`Two-Stage Model`, y=`Joint Model`), size=0.5) + geom_abline() +
  scale_x_continuous(trans='log10', labels = label_comma()) + scale_y_continuous(trans='log10', labels = label_comma()) +
   coord_cartesian(ylim=c(0.03,0.5), xlim=c(0.03,0.5)) + facet_wrap( ~ validation, ncol = 2) #+
  #ggtitle('Species validation RMSE: Certain vs. Uncertain Abiotics')
ggsave('results_paper/rmse_comparison.png', width=1200, heigh=800, units='px')

```

Compare predicted and observed prevalence in each province:

```{r}
prevalences <- read.csv("data_paper/prevalences_province.csv")
# prevalences <- read.csv("data_paper/prevalences_province.csv", col.names=c("area_prevalences", "meta", "model", "species", "naam")) # LSF cluster results

prevalences.predicted <- prevalences %>% group_by(meta, species, naam) %>% 
  summarize(mean = mean(area_prevalences, na.rm=T), 
            X0.025quant = quantile(area_prevalences, 0.025, na.rm=T),
            X0.975quant = quantile(area_prevalences, 0.975, na.rm=T))

prevalences.observed <- sf.plots %>% as.data.frame  %>% select(naam, all_of(species)) %>% 
  #mutate(naam=Fys_geo_regio) %>% filter(!(naam %in% c("gtw", "gtz"))) %>% 
  group_by(naam) %>% summarize_all(function(x) mean(x, na.rm=T)) %>% 
  pivot_longer(species, names_to = "species", values_to="mean") %>% mutate(meta="True")

prevalences.both <- rbind(prevalences.predicted, prevalences.observed) %>%
  filter((species != "Struikhei") & (!is.na(naam))) %>%
  mutate(naam=recode(naam, "Drenthe"='DR', "Flevoland"='FL', "Fryslân"='FR', "Gelderland"='GE', "Groningen"='GR',
                     "Limburg"='LI', "Noord-Brabant"='NB', "Noord-Holland"='NH', "Overijssel"='OV', "Utrecht"='UT',
                     "Zeeland"='ZE', "Zuid-Holland"='ZH')) %>%
  mutate(meta=factor(recode(meta, "Certain" = "Two-Stage", "Uncertain"="Joint"), levels=c("Two-Stage", "True", "Joint"))) %>%
  merge(species.map) %>% mutate(species = name)

plot.all <- ggplot(prevalences.both, aes(x=naam, y=mean, col=meta, group=meta)) + 
  geom_point(position=position_dodge(width=0.5), size=2) + 
  geom_errorbar(position=position_dodge(width=0.5), width=0.5, aes(ymin=X0.025quant, ymax=X0.975quant)) +
  theme(legend.position="top", axis.text.x = element_text(angle = 90)) + #+ position_dodge()
  scale_color_discrete(name = "Model") +
  scale_x_discrete(name='Province') + 
  scale_y_continuous(name='Prevalence') +
  facet_wrap( ~ species, ncol = 5, scales='free_y')
plot.all
ggsave('results_paper/all_prevalences.png', width=2800, heigh=4000, units='px')
```

Compare predicted and observed prevalence in each FGR region:

```{r}
prevalences <- read.csv("data_paper/prevalences_fgr.csv")
# prevalences <- read.csv("data_paper/prevalences_fgr.csv", col.names=c("area_prevalences", "meta", "model", "species", "naam")) # LSF cluster results

prevalences.predicted <- prevalences %>% group_by(meta, species, naam) %>% 
  summarize(mean = mean(area_prevalences, na.rm=T), 
            X0.025quant = quantile(area_prevalences, 0.025, na.rm=T),
            X0.975quant = quantile(area_prevalences, 0.975, na.rm=T))

prevalences.observed <- sf.plots %>% as.data.frame  %>% select(Fys_geo_regio, all_of(species)) %>% 
  mutate(naam=Fys_geo_regio) %>% filter(!(naam %in% c("gtw", "gtz"))) %>% 
  group_by(naam) %>% summarize_all(function(x) mean(x, na.rm=T)) %>% 
  pivot_longer(species, names_to = "species", values_to="mean") %>% mutate(meta="True")

prevalences.both <- rbind(prevalences.predicted, prevalences.observed) %>%
  filter((species != "Struikhei") & (!is.na(naam))) %>%
  mutate(meta=factor(recode(meta, "Certain" = "Two-Stage", "Uncertain"="Joint"), levels=c("Two-Stage", "True", "Joint")))  %>%
  merge(species.map) %>% mutate(species = name) #%>%
  #mutate(naam=recode(naam, "Drenthe"='DR', "Flevoland"='FL', "Fryslân"='FR', "Gelderland"='GE', "Groningen"='GR',
  #                   "Limburg"='LI', "Noord-Brabant"='NB', "Noord-Holland"='NH', "Overijssel"='OV', "Utrecht"='UT',
  #                   "Zeeland"='ZE', "Zuid-Holland"='ZH'))

plot.all <- ggplot(prevalences.both, aes(x=naam, y=mean, col=meta, group=meta)) + 
  geom_point(position=position_dodge(width=0.5), size=2) + 
  geom_errorbar(position=position_dodge(width=0.5), width=0.5, aes(ymin=X0.025quant, ymax=X0.975quant)) +
  theme(legend.position="top", axis.text.x = element_text(angle = 90)) + #+ position_dodge()
  scale_color_discrete(name = "Model") +
  scale_x_discrete(name='Fys_geo_regio') + 
  scale_y_continuous(name='Prevalence') +
  facet_wrap( ~ species, ncol = 5, scales='free_y')
plot.all
ggsave('results_paper/all_prevalences2.png', width=2800, heigh=4000, units='px')
```

### 2023-08-15 paper_data.Rmd
Add files via upload

---
title: "paper_data"
author: "Markus Viljanen"
date: "2023-08-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
library(readxl)
library(RPostgreSQL)
library(plyr)
library(tidyverse)
library(sf)
```

## Original LMF data set to LMF subset

Fetch Netherlands boundaries from RIVM GIS server

```{r}
conn <- dbConnect(PostgreSQL(), dbname="geodata", user = "ro", password="ro!geo", host = "geodatabase.rivm.nl")
NL <- st_read(conn, query = "SELECT * FROM nl.adm_wijkbuurt_2022_gen;") %>%
  filter(indeling == 'gemeente') %>% st_union()
st_write(NL, "data_paper/NL.shp", append=F)
```

Abiotic measurements for LMF data

```{r}
# Abiotic measurements for LMF data 
LMF.dir <- file.path('LMF', 'LMF data.xlsx')
df.abiotic <- read_excel(LMF.dir, 1)[-1,-1] %>% distinct(releve_nr, .keep_all=T) %>% 
  mutate(across(DATE:`N-(NO3+NO2)`,as.numeric)) %>% mutate(NAAM=tolower(str_replace_all(NAAM, " ", "")))
# Fix some identifier names in abiotic data
naam <- df.abiotic$NAAM
# some FR**** and FR**** identifiers were not properly zero padded
replace.index <- substr(naam,1,2) %in% c("fr", "fl")
pad.4 <- paste(substr(naam,1,2), str_pad(substr(naam,3,nchar(naam)), 4, pad="0"), sep="")
df.abiotic[replace.index, "NAAM"] <- pad.4[replace.index]
# Some GR**** identifiers were GR****/GR****
replace.index <- grepl('/', df.abiotic$NAAM, fixed = TRUE)
last.6 <- substr(naam, nchar(naam)-6+1, nchar(naam))
df.abiotic[replace.index, "NAAM"] <- last.6[replace.index]
# Lisa's fixes to coordinates 
# Tostrams, Lisa <lisa.tostrams@rivm.nl>
df.abiotic.2 <- read.csv('LMF/lmf_log_data_corrected_coords.csv') %>% 
  distinct(releve_nr, .keep_all=T) %>% 
  select(releve_nr, `X.coördinaat`, `Y.coördinaat`)
# Replace coordinates
df.abiotic <- df.abiotic %>% left_join(df.abiotic.2)  %>% 
  mutate(Plot = NAAM %>% factor, X = `X.coördinaat`, Y=`Y.coördinaat`) %>%
  group_by(Plot) %>% filter(DATE == max(DATE)) %>% filter(releve_nr != 203857) #NAAM = nb1019 has two releve_nr in same year
```

Original LMF data (Species names, Field visit information, Vegatation plots information, Species observations)

```{r}
# Original LMF data
# https://longreads.cbs.nl/meetprogrammas-flora-en-fauna-2022/

# Species
df.species = read.csv('LMF/LMF_Soorten.csv',sep=';', colClasses=c("numeric", "numeric", "factor", "factor")) %>% 
  select(Soortcode, Nednaam, Wetnaam)

# Visits
df.visit <- read.csv('LMF/LMF_kop.csv', sep=';', colClasses=c("factor", "numeric", "character", "numeric", "numeric", 
                                                              "factor", "factor", "factor", "factor", "factor", 
                                                              "factor", "factor", "factor", "factor", 
                                                              "numeric", "numeric", "numeric", "factor", "factor", "factor")) %>% 
  mutate(Datum=as.Date(Datum, format = "%Y%m%d")) %>% 
  select(Plot, Jaar, Dagnummer, Pq_grootte)

# Plots
df.plots <- read.csv('LMF/LMF_plot.csv',sep=';', dec=',', colClasses=c("factor", "numeric", "numeric", 
                                                                       "factor", "factor", "factor", "factor", "factor", 
                                                                       "factor", "factor", "factor", "factor", "numeric", "factor")) %>%
  mutate(X=X*1000, Y=Y*1000) %>%
  #mutate(x = as.numeric(gsub(",", ".", X))*1000,
  #       y = as.numeric(gsub(",", ".", Y))*1000) %>%
  select(Plot, X, Y, Fys_geo_regio)

# Observations, take only latest
df.occurrence <- read.csv('LMF/LMF_tellingen.csv', sep=';', colClasses=c("factor", "numeric", "numeric", "numeric")) %>% 
  select(Plot, Jaar, Soortcode, Bedekking)
df.occurrence <- df.occurrence %>% merge(df.species) %>% group_by(Plot) %>% filter(Jaar == max(Jaar))

# Add plot visit date 
df.extra <- df.occurrence %>% distinct(Plot, Jaar) %>% merge(df.visit)
df.plots <- df.plots %>% merge(df.extra)

# Add abiotic information
df.plots <- df.plots %>% full_join(df.abiotic, by="Plot", suffix = c("", ".abiotic")) %>%
  mutate(X=ifelse(is.na(X), X.abiotic, X),
         Y=ifelse(is.na(Y), Y.abiotic, Y))
```

Wieger's proposed species to analyze

```{r}
# Wamelink, Wieger <wieger.wamelink@wur.nl>
df.n <- df.occurrence %>% group_by(Nednaam) %>% count() %>% arrange(-n)
LMF.dir <- file.path('LMF', '8969892_WageningenEnvironmentalResearch_6708PB_20201006_ww3.xlsx')
species_proposal <- c(read_excel(LMF.dir)$NedNAAM, "Gestreepte witbol", "Struikhei")
species <- (data.frame(Nednaam=species_proposal) %>% merge(df.n) %>% filter(n>=20) %>% select(Nednaam))$Nednaam
chosen.species <- data.frame(Nednaam=species[species != "Struikhei"]) %>% 
  merge(df.species) %>% merge(df.n) %>% select(Nednaam, Wetnaam, n) %>% arrange(-n) 
```

Overlap statistics:

```{r}
c1 <- unique(df.abiotic$Plot)
c2 <- unique(df.visit$Plot)
cx <- intersect(c1, c2)
sprintf("Plots: %d abiotic, %d visit, %d intersection", length(c1), length(c2), length(cx))

c1 <- unique(df.abiotic$DATE)
c2 <- unique(df.visit$Jaar)
cx <- intersect(c1, c2)
sprintf("Years: %d abiotic, %d visit, %d intersection", length(c1), length(c2), length(cx))

c1 <- unique(paste(as.character(df.abiotic$Plot), 'x', as.character(df.abiotic$DATE)))
c2 <- unique(paste(as.character(df.visit$Plot), 'x', as.character(df.visit$Jaar)))
cx <- intersect(c1, c2)
sprintf("(Plot,Year)-pairs: %d abiotic, %d visit, %d intersection", length(c1), length(c2), length(cx))
```

Binary table of each species occurrence at each vegetation plot (0/1)

```{r}
df.y <- df.occurrence %>% select(Plot, Nednaam, Bedekking) %>% mutate(Bedekking=1) %>% spread(Nednaam, Bedekking, fill=0)
```

Take only relevant vegetation plot information and join to species occurrence at each plot

```{r}
df.plots <- df.plots %>% 
  select(Plot, releve_nr, X, Y, Fys_geo_regio, Jaar, Dagnummer, Pq_grootte, `pH-H20`, ORG_STOF, `C/N`, N_totaal, P_totaal, K_CaCl2) %>% 
  mutate(across(`pH-H20`:K_CaCl2, function (x) round(x, digits=2))) %>%
  full_join(df.y[c("Plot", species)])
```

## Add information, spatial joins with geodata

In codes we use X&Y as km to enforce a vague prior

```{r}
NL <- (NL - matrix(data=c(155000, 463000), ncol=2)) * matrix(data=c(1/1000, 1/1000), ncol=2)
```

Data set of species observations and environmental conditions

```{r}
sf.plots <- df.plots %>% st_as_sf(coords = c("X","Y"), crs=st_crs(NL))
st_geometry(sf.plots) <- (sf.plots$geometry - matrix(data=c(155000, 463000), ncol=2)) * matrix(data=c(1/1000, 1/1000), ncol=2)
sf.plots[c("X", "Y")] <- sf.plots  %>% st_coordinates()
```

Grid of all possible plots

```{r}
# Grid with rectangles and their centers  #%>% st_as_sf(crs=st_crs(NL))
grid.rect <- st_make_grid(NL,n=100,crs=st_crs(NL), what='polygons', square=T) 
grid.circle <- st_make_grid(NL,n=100,crs=st_crs(NL), what='centers')
index <- which(lengths(st_within(grid.circle, NL)) > 0) #st_intersects(grid.rect, NL)
# Create data frame
sf.grid <- data.frame(grid.circle[index], grid.rect[index]) %>% st_as_sf(crs=st_crs(NL))
sf.grid[c("X", "Y")] <- sf.grid %>% st_geometry %>% st_centroid() %>% st_coordinates()
sf.grid$Plot <- factor(sprintf("Grid%d", 1:nrow(sf.grid)))
```

Add GIS layers: province

```{r}
#https://service.pdok.nl/kadaster/bestuurlijkegebieden/atom/v1_0/bestuurlijke_gebieden.xml
provincie = st_read("BestuurlijkeGebieden_2023.gpkg", "provinciegebied") %>% select(naam)
st_geometry(provincie) <- (provincie$geom - matrix(data=c(155000, 463000), ncol=2)) * matrix(data=c(1/1000, 1/1000), ncol=2)

sf.plots <- sf.plots %>% st_join(provincie , left=T, st_within)
sf.grid <- sf.grid %>% st_join(provincie , left=T, st_within)
#plot(provincie)
```

Add GIS layers: FGR regio

```{r}
# File created by FGR.ipynb
fgr = st_read("FGR_lmf_mapping.gpkg") %>% select(lmf_indeling)
st_geometry(fgr) <- (fgr$geom - matrix(data=c(155000, 463000), ncol=2)) * matrix(data=c(1/1000, 1/1000), ncol=2)

sf.plots <- sf.plots %>% st_join(fgr , left=T, st_within)
sf.grid <- sf.grid %>% st_join(fgr , left=T, st_within)
#plot(fgr)
```

Add GIS layers: soiltype, landuse

```{r}
# https://www.nationaalgeoregister.nl/geonetwork/srv/dut/catalog.search#/metadata/2d3dd6d2-2d2b-4b5f-9e30-86e19ed77a56
# http://www.geodesk.nl/Grondsoorten.htm
## takes ~ 5 mins
landuse <- st_read("bestand_bodemgebruik_2015.gpkg", layer="bestand_bodemgebruik_2015")
soiltype <- st_read("Grondsoortenkaart-2006.shp")
landuse <- landuse %>% transmute(bodemgebruik=relevel(factor(bodemgebruik), ref="Bos"))
soiltype <- soiltype %>% transmute(OMSCHRIJVI=relevel(factor(OMSCHRIJVI), ref="Zand"))
st_geometry(landuse) <- (landuse$geom - matrix(data=c(155000, 463000), ncol=2)) * matrix(data=c(1/1000, 1/1000), ncol=2)
st_geometry(soiltype) <- (soiltype$geometry - matrix(data=c(155000, 463000), ncol=2)) * matrix(data=c(1/1000, 1/1000), ncol=2)
```

```{r}
## takes ~2 hours
info.plots <- sf.plots %>% select(Plot)
info.plots <- info.plots %>% st_join(landuse , left=T, st_within)
info.plots <- info.plots %>% st_join(soiltype , left=T, st_within)
info.plots <- info.plots %>% as.data.frame %>% distinct(Plot, bodemgebruik, OMSCHRIJVI)
#info.plots %>% saveRDS("plot.info.RDS")
#info.plots <- readRDS("plot.info.RDS")
sf.plots <- sf.plots %>% merge(info.plots, on="Plot")
```

```{r}
## takes ~2 hours
info.grid <- sf.grid %>% select(Plot)
info.grid <- info.grid %>% st_join(landuse , left=T, st_within)
info.grid <- info.grid %>% st_join(soiltype , left=T, st_within)
info.grid <- info.grid %>% as.data.frame %>% distinct(Plot, bodemgebruik, OMSCHRIJVI)
#info.grid %>% saveRDS("grid.info.RDS")
#info.grid <- readRDS("grid.info.RDS")
# sf.grid %>% merge(info.grid, on="Plot") drops the second geometry, so cbind the correctly sorted auxiliary data
sf.grid <- cbind(sf.grid, sf.grid %>% as.data.frame %>% select(Plot) %>% merge(info.grid, on="Plot", sort=F) %>% select(bodemgebruik, OMSCHRIJVI)) 
```

Fill very few NA observations with most common category and combine some categories

```{r}
sf.plots <- sf.plots %>% mutate(
  bodemgebruik= fct_na_value_to_level(fct_collapse(bodemgebruik, 
      Bebouwd=c("Bebouwd exclusief bedrijfsterrein", "Bedrijfsterrein", "Semi-bebouwd", "Glastuinbouw"),
      Spoorterrein=c("Spoorterrein", "Vliegveld")), "Landbouw en overig agrarisch"),
 OMSCHRIJVI=fct_na_value_to_level(relevel(factor(OMSCHRIJVI), ref="Zand"), "Zand"))

sf.grid <- sf.grid %>% mutate(
  bodemgebruik=fct_na_value_to_level(fct_collapse(bodemgebruik, 
      Bebouwd=c("Bebouwd exclusief bedrijfsterrein", "Bedrijfsterrein", "Semi-bebouwd", "Glastuinbouw"),
      Spoorterrein=c("Spoorterrein", "Vliegveld")), "Landbouw en overig agrarisch"),
  OMSCHRIJVI=fct_na_value_to_level(relevel(factor(OMSCHRIJVI), ref="Zand"), "Zand"))
```

Save data

```{r}
chosen.species %>% write.csv(file="data_paper/chosen_species.csv", row.names=F)
sf.plots %>% as.data.frame %>% select(-geometry) %>% write.csv(file="data_paper/data_plots.csv", row.names=F)
sf.grid %>% as.data.frame %>% select(-geometry,-geometry.1) %>% write.csv(file="data_paper/data_grid.csv", row.names=F)
```

### 2023-08-15 lsf/predict_uncertain.r

@@ -1,0 +1,131 @@
+args <- commandArgs(trailingOnly = TRUE)
+species.take <- args[[1]]
+provincie <- args[[2]]
+print(species.take)
+print(provincie)
+
+library(INLA)
+library(dplyr)
+library(stringr)
+library(tibble)
+
+sf.plots <- readRDS("sf.plots.RDS")
+sf.grid <- readRDS("sf.grid.RDS")
+
+xy.plots <- sf.plots %>% as.data.frame %>% select(X,Y)
+xy.grid <- sf.grid %>% as.data.frame %>% select(X,Y)
+
+step <- 10000 / 1000
+domain <- inla.nonconvex.hull(as.matrix(xy.plots), convex=-0.04)
+mesh <- inla.mesh.2d(boundary=domain, max.edge=c(step*2, step*4), cutoff=step)
+
+spde <- inla.spde2.matern(mesh=mesh,alpha=2)##print(spde$n.spde)
+s.index <- inla.spde.make.index(name="spatial.field", n.spde=spde$n.spde) 
+
+A_y <- inla.spde.make.A(mesh=mesh, loc=as.matrix(xy.plots))
+A_pred <- inla.spde.make.A(mesh=mesh, loc=as.matrix(xy.grid))
+variables <- c("ORG_STOF.z", "CN.z", "N_totaal.z", "P_totaal.z", "K_CaCl2.z")
+
+i <- 1
+stacks <- list()
+for (var in variables) {
+# Observations of the variable
+sf.var <- sf.plots %>% filter(!is.na(get(var)))
+A_var <- inla.spde.make.A(mesh=mesh, loc=as.matrix(sf.var %>% as.data.frame %>% select(X,Y)))
+# One output for every estimated variable (+ species occurrence)
+y <- matrix(nrow=nrow(sf.var), ncol=length(variables) + 1)
+y.0 <- matrix(nrow=nrow(sf.grid), ncol=length(variables) + 1)
+y[,i] <- sf.var[[var]]
+# Stack
+effects <- list()
+effects[[sprintf("%s.i.field", var)]] = 1:spde$n.spde
+effects[[sprintf("%s.intercept", var)]] = 1 #asdf
+stacks[[var]] <- inla.stack(data=list(y=y), effects=list(effects), A=list(A_var), tag=sprintf("est.%s", var))
[…]

### 2023-08-15 lsf/results_abiotic.r

@@ -1,0 +1,96 @@
+args <- commandArgs(trailingOnly = TRUE)
+
+library(INLA)
+library(dplyr)
+library(stringr)
+library(tibble)
+
+sf.plots <- readRDS("sf.plots.RDS")
+sf.grid <- readRDS("sf.grid.RDS")
+
+xy.plots <- sf.plots %>% as.data.frame %>% select(X,Y)
+xy.grid <- sf.grid %>% as.data.frame %>% select(X,Y)
+
+step <- 10000 / 1000
+domain <- inla.nonconvex.hull(as.matrix(xy.plots), convex=-0.04)
+mesh <- inla.mesh.2d(boundary=domain, max.edge=c(step*2, step*4), cutoff=step)
+
+spde <- inla.spde2.matern(mesh=mesh,alpha=2)##print(spde$n.spde)
+s.index <- inla.spde.make.index(name="spatial.field", n.spde=spde$n.spde) 
+
+A_y <- inla.spde.make.A(mesh=mesh, loc=as.matrix(xy.plots))
+A_pred <- inla.spde.make.A(mesh=mesh, loc=as.matrix(xy.grid))
+variables <- c("ORG_STOF.z", "CN.z", "N_totaal.z", "P_totaal.z", "K_CaCl2.z")
+
+
+
+terms.abiotics <- list()
+i <- 1
+for (variable in variables) {
+  print(variable)
+
+  stack.est <- inla.stack(data=list(y = sf.plots[[variable]]),
+                          A=list(A_y), effects=list(c(s.index, list(intercept=1))), 
+                          tag="est")
+  stack.pred <- inla.stack(data=list(y=NA), 
+                           A=list(A_pred), effects=list(c(s.index, list(intercept=1))),
+                           tag="pred") 
+  join.stack <- inla.stack(stack.est, stack.pred)
+  
+  formula <- y ~ -1 + intercept + f(spatial.field, model=spde)
+  
+  output <- inla(formula, family="gaussian", num.threads=1,
+                      data=inla.stack.data(join.stack, spde=spde),
+                      control.predictor=list(A=inla.stack.A(join.stack), compute=TRUE, link=1),
+                control.inla = list(strategy = "adaptive", int.strategy = "eb"))
+  
+  # predictions at field visits
+  index.est <- inla.stack.index(join.stack, tag="est")$data
+  variable.est <- output$summary.fitted.values[index.est,]
[…]

### 2023-08-15 lsf/results_certain.r

@@ -1,0 +1,111 @@
+args <- commandArgs(trailingOnly = TRUE)
+species.take <- args[[1]]
+print(species.take)
+
+library(INLA)
+library(dplyr)
+library(stringr)
+library(tibble)
+
+sf.plots <- readRDS("sf.plots.RDS")
+sf.grid <- readRDS("sf.grid.RDS")
+
+xy.plots <- sf.plots %>% as.data.frame %>% select(X,Y)
+xy.grid <- sf.grid %>% as.data.frame %>% select(X,Y)
+
+step <- 10000 / 1000
+domain <- inla.nonconvex.hull(as.matrix(xy.plots), convex=-0.04)
+mesh <- inla.mesh.2d(boundary=domain, max.edge=c(step*2, step*4), cutoff=step)
+
+spde <- inla.spde2.matern(mesh=mesh,alpha=2)##print(spde$n.spde)
+s.index <- inla.spde.make.index(name="spatial.field", n.spde=spde$n.spde) 
+
+A_y <- inla.spde.make.A(mesh=mesh, loc=as.matrix(xy.plots))
+A_pred <- inla.spde.make.A(mesh=mesh, loc=as.matrix(xy.grid))
+variables <- c("ORG_STOF.z", "CN.z", "N_totaal.z", "P_totaal.z", "K_CaCl2.z")
+
+
+stk_y <- inla.stack(data = list(y=sf.plots[[species.take]]),
+              A = list(A_y, 1, 1, 1, 1, 1, 1, 1, 1), 
+              effects = list(c(list(u.field=1:spde$n.spde), list(intercept=1)), 
+                              list(terrain=sf.plots$terrain), 
+                              list(Dagnummer.z=sf.plots$Dagnummer.z), 
+                              list(pH.z=sf.plots$pH.z), 
+                              list(ORG_STOF.z=sf.plots$ORG_STOF.z.mean),
+                              list(CN.z=sf.plots$CN.z.mean), 
+                              list(N_totaal.z=sf.plots$N_totaal.z.mean), 
+                              list(P_totaal.z=sf.plots$P_totaal.z.mean), 
+                              list(K_CaCl2.z=sf.plots$K_CaCl2.z.mean)
+                              ),
+              tag="est.y")
+
+n <- nrow(sf.grid)
+stk_pred <- inla.stack(data=list(y=rep(NA, n)),  
+                        effects=list(c(list(u.field=1:spde$n.spde), list(intercept=1)),
+                                      list(terrain=sf.grid$terrain), 
[…]

### 2023-08-15 lsf/results_uncertain.r

@@ -1,0 +1,195 @@
+args <- commandArgs(trailingOnly = TRUE)
+species.take <- args[[1]]
+
+print(species.take)
+
+library(INLA)
+library(dplyr)
+library(stringr)
+library(tibble)
+
+sf.plots <- readRDS("sf.plots.RDS")
+sf.grid <- readRDS("sf.grid.RDS")
+
+xy.plots <- sf.plots %>% as.data.frame %>% select(X,Y)
+xy.grid <- sf.grid %>% as.data.frame %>% select(X,Y)
+
+step <- 10000 / 1000
+domain <- inla.nonconvex.hull(as.matrix(xy.plots), convex=-0.04)
+mesh <- inla.mesh.2d(boundary=domain, max.edge=c(step*2, step*4), cutoff=step)
+
+spde <- inla.spde2.matern(mesh=mesh,alpha=2)##print(spde$n.spde)
+s.index <- inla.spde.make.index(name="spatial.field", n.spde=spde$n.spde) 
+
+A_y <- inla.spde.make.A(mesh=mesh, loc=as.matrix(xy.plots))
+A_pred <- inla.spde.make.A(mesh=mesh, loc=as.matrix(xy.grid))
+variables <- c("ORG_STOF.z", "CN.z", "N_totaal.z", "P_totaal.z", "K_CaCl2.z")
+
+i <- 1
+stacks <- list()
+for (var in variables) {
+  # Observations of the variable
+  sf.var <- sf.plots %>% filter(!is.na(get(var)))
+  A_var <- inla.spde.make.A(mesh=mesh, loc=as.matrix(sf.var %>% as.data.frame %>% select(X,Y)))
+  # One output for every estimated variable (+ species occurrence)
+  y <- matrix(nrow=nrow(sf.var), ncol=length(variables) + 1)
+  y.0 <- matrix(nrow=nrow(sf.grid), ncol=length(variables) + 1)
+  y[,i] <- sf.var[[var]]
+  # Stack
+  effects <- list()
+  effects[[sprintf("%s.i.field", var)]] = 1:spde$n.spde
+  effects[[sprintf("%s.intercept", var)]] = 1 #asdf
+  stacks[[var]] <- inla.stack(data=list(y=y), effects=list(effects), A=list(A_var), tag=sprintf("est.%s", var))
[…]

### 2023-08-18 paper_data.Rmd
some review

@@
 ## Original LMF data set to LMF subset
-
-Fetch Netherlands boundaries from RIVM GIS server
-
-```{r}
-conn <- dbConnect(PostgreSQL(), dbname="geodata", user = "ro", password="ro!geo", host = "geodatabase.rivm.nl")
-NL <- st_read(conn, query = "SELECT * FROM nl.adm_wijkbuurt_2022_gen;") %>%
-  filter(indeling == 'gemeente') %>% st_union()
-st_write(NL, "data_paper/NL.shp", append=F)
-```
+Here as well as in the modelling code, all coordinates are in EPSG:28992 Projected coordinate system for Netherlands.
 
@@
 # Lisa's fixes to coordinates 
+# This consisted of 24 plots of which the locations seemed to be outside of the Netherlands.
+The plotnames were found in the LMF plots, and for those the correct coordinates were taken from the LMF programme.
+# There were also an additional 329 abiotic plots with a plot name that was in the LMF file, but where the coordinates did not match.
+In those cases the coordinates from the LMF programme were taken as well.
 # Tostrams, Lisa <lisa.tostrams@rivm.nl>
@@
 ```{r}
+NL <- st_read("data_paper/NL.shp")$geometry
 NL <- (NL - matrix(data=c(155000, 463000), ncol=2)) * matrix(data=c(1/1000, 1/1000), ncol=2)

### 2023-08-23 paper_codes.Rmd
final commit

@@
 ## Libraries
+For INLA, see: https://www.r-inla.org/download-install
 
@@
 Netherlands boundary shapefile
+Coordinates are in EPSG:28992, the Projected coordinate system for Netherlands.
+These coordinates are calculated from the center of the Netherlands in meters.
+For simplicity, we convert them to kilometers below.
 
@@
 Data for vegetation plots and abiotic measurements data
+Coordinates are in EPSG:28992, the Projected coordinate system for Netherlands.
 
@@
 Data for Netherlands grid
+Coordinates are in EPSG:28992, the Projected coordinate system for Netherlands.
 
@@
 ## Experiments: training and validation
+It is also possible to download the results from https://zenodo.org/record/8261504
+In that case, place the downloaded files in the subfolder `data_paper/` and go to the `Results: import file` section below to import the results and create the plots.
 
@@
 Combine and export the results of fitting all models to data: 
+It is also possible to skip this cell and download the results from https://zenodo.org/record/8261504
+Place the data in the subfolder `data_paper/` and run the code below this cell to create the plots.
 
@@
 
-## Results: crowberry example
+## Results: import file
 
@@
 ```
+
+## Results: crowberry example
 

### 2023-08-23 paper_data.Rmd
final commit

@@
 ## Original LMF data set to LMF subset
-Here as well as in the modelling code, all coordinates are in EPSG:28992 Projected coordinate system for Netherlands.
+Fetch Netherlands boundaries from RIVM GIS server. Here as well as in the modelling code, all coordinates are in EPSG:28992 Projected coordinate system for Netherlands.
+
+```{r}
+conn <- dbConnect(PostgreSQL(), dbname="geodata", user = "", password="", host = "geodatabase.rivm.nl")
+NL <- st_read(conn, query = "SELECT * FROM nl.adm_wijkbuurt_2022_gen;") %>%
+  filter(indeling == 'gemeente') %>% st_union()
+st_write(NL, "data_paper/NL.shp", append=F)
+```
 
@@
 ```{r}
-NL <- st_read("data_paper/NL.shp")$geometry
 NL <- (NL - matrix(data=c(155000, 463000), ncol=2)) * matrix(data=c(1/1000, 1/1000), ncol=2)

### 2023-08-23 lsf/predict_certain.r

@@ -1,0 +1,104 @@
+args <- commandArgs(trailingOnly = TRUE)
+species.take <- args[[1]]
+provincie <- args[[2]]
+print(species.take)
+print(provincie)
+
+library(INLA)
+library(dplyr)
+library(stringr)
+library(tibble)
+
+sf.plots <- readRDS("sf.plots.RDS")
+sf.grid <- readRDS("sf.grid.RDS")
+
+xy.plots <- sf.plots %>% as.data.frame %>% select(X,Y)
+xy.grid <- sf.grid %>% as.data.frame %>% select(X,Y)
+
+step <- 10000 / 1000
+domain <- inla.nonconvex.hull(as.matrix(xy.plots), convex=-0.04)
+mesh <- inla.mesh.2d(boundary=domain, max.edge=c(step*2, step*4), cutoff=step)
+
+spde <- inla.spde2.matern(mesh=mesh,alpha=2)##print(spde$n.spde)
+s.index <- inla.spde.make.index(name="spatial.field", n.spde=spde$n.spde) 
+
+A_y <- inla.spde.make.A(mesh=mesh, loc=as.matrix(xy.plots))
+A_pred <- inla.spde.make.A(mesh=mesh, loc=as.matrix(xy.grid))
+variables <- c("ORG_STOF.z", "CN.z", "N_totaal.z", "P_totaal.z", "K_CaCl2.z")
+
+
+sf.train = sf.plots %>% filter(naam != provincie) #%>% filter(bodemgebruik %in% c("Bos", "Droog natuurlijk terrein", "Nat natuurlijk terrein"))
+sf.test = sf.plots %>% filter(naam == provincie) #%>% filter(bodemgebruik %in% c("Bos", "Droog natuurlijk terrein", "Nat natuurlijk terrein"))
+A_train <- inla.spde.make.A(mesh=mesh, loc=as.matrix(sf.train %>% as.data.frame %>% select(X,Y)))
+A_test <- inla.spde.make.A(mesh=mesh, loc=as.matrix(sf.test %>% as.data.frame %>% select(X,Y)))
+
+### Certain abiotics ###
+
+# Fit
+stk_y <- inla.stack(data = list(y=sf.train[[species.take]]),
+                A = list(A_train, 1, 1, 1, 1, 1, 1, 1, 1), 
+                effects = list(c(list(u.field=1:spde$n.spde), list(intercept=1)), 
+                                list(terrain=sf.train$terrain), 
[…]

### 2023-08-23 lsf/predict_certain2.r

@@ -1,0 +1,104 @@
+args <- commandArgs(trailingOnly = TRUE)
+species.take <- args[[1]]
+geo <- args[[2]]
+print(species.take)
+print(geo)
+
+library(INLA)
+library(dplyr)
+library(stringr)
+library(tibble)
+
+sf.plots <- readRDS("sf.plots.RDS")
+sf.grid <- readRDS("sf.grid.RDS")
+
+xy.plots <- sf.plots %>% as.data.frame %>% select(X,Y)
+xy.grid <- sf.grid %>% as.data.frame %>% select(X,Y)
+
+step <- 10000 / 1000
+domain <- inla.nonconvex.hull(as.matrix(xy.plots), convex=-0.04)
+mesh <- inla.mesh.2d(boundary=domain, max.edge=c(step*2, step*4), cutoff=step)
+
+spde <- inla.spde2.matern(mesh=mesh,alpha=2)##print(spde$n.spde)
+s.index <- inla.spde.make.index(name="spatial.field", n.spde=spde$n.spde) 
+
+A_y <- inla.spde.make.A(mesh=mesh, loc=as.matrix(xy.plots))
+A_pred <- inla.spde.make.A(mesh=mesh, loc=as.matrix(xy.grid))
+variables <- c("ORG_STOF.z", "CN.z", "N_totaal.z", "P_totaal.z", "K_CaCl2.z")
+
+
+sf.train = sf.plots %>% filter(Fys_geo_regio != geo) #%>% filter(bodemgebruik %in% c("Bos", "Droog natuurlijk terrein", "Nat natuurlijk terrein"))
+sf.test = sf.plots %>% filter(Fys_geo_regio == geo) #%>% filter(bodemgebruik %in% c("Bos", "Droog natuurlijk terrein", "Nat natuurlijk terrein"))
+A_train <- inla.spde.make.A(mesh=mesh, loc=as.matrix(sf.train %>% as.data.frame %>% select(X,Y)))
+A_test <- inla.spde.make.A(mesh=mesh, loc=as.matrix(sf.test %>% as.data.frame %>% select(X,Y)))
+
+### Certain abiotics ###
+
+# Fit
+stk_y <- inla.stack(data = list(y=sf.train[[species.take]]),
+                A = list(A_train, 1, 1, 1, 1, 1, 1, 1, 1), 
+                effects = list(c(list(u.field=1:spde$n.spde), list(intercept=1)), 
+                                list(terrain=sf.train$terrain), 
[…]

### 2023-08-23 lsf/predict_uncertain2.r

@@ -1,0 +1,132 @@
+args <- commandArgs(trailingOnly = TRUE)
+species.take <- args[[1]]
+geo <- args[[2]]
+print(species.take)
+print(geo)
+
+
+library(INLA)
+library(dplyr)
+library(stringr)
+library(tibble)
+
+sf.plots <- readRDS("sf.plots.RDS")
+sf.grid <- readRDS("sf.grid.RDS")
+
+xy.plots <- sf.plots %>% as.data.frame %>% select(X,Y)
+xy.grid <- sf.grid %>% as.data.frame %>% select(X,Y)
+
+step <- 10000 / 1000
+domain <- inla.nonconvex.hull(as.matrix(xy.plots), convex=-0.04)
+mesh <- inla.mesh.2d(boundary=domain, max.edge=c(step*2, step*4), cutoff=step)
+
+spde <- inla.spde2.matern(mesh=mesh,alpha=2)##print(spde$n.spde)
+s.index <- inla.spde.make.index(name="spatial.field", n.spde=spde$n.spde) 
+
+A_y <- inla.spde.make.A(mesh=mesh, loc=as.matrix(xy.plots))
+A_pred <- inla.spde.make.A(mesh=mesh, loc=as.matrix(xy.grid))
+variables <- c("ORG_STOF.z", "CN.z", "N_totaal.z", "P_totaal.z", "K_CaCl2.z")
+
+i <- 1
+stacks <- list()
+for (var in variables) {
+# Observations of the variable
+sf.var <- sf.plots %>% filter(!is.na(get(var)))
+A_var <- inla.spde.make.A(mesh=mesh, loc=as.matrix(sf.var %>% as.data.frame %>% select(X,Y)))
+# One output for every estimated variable (+ species occurrence)
+y <- matrix(nrow=nrow(sf.var), ncol=length(variables) + 1)
+y.0 <- matrix(nrow=nrow(sf.grid), ncol=length(variables) + 1)
+y[,i] <- sf.var[[var]]
+# Stack
+effects <- list()
+effects[[sprintf("%s.i.field", var)]] = 1:spde$n.spde
+effects[[sprintf("%s.intercept", var)]] = 1 #asdf
+stacks[[var]] <- inla.stack(data=list(y=y), effects=list(effects), A=list(A_var), tag=sprintf("est.%s", var))
[…]

### 2023-08-23 lsf/run.sh

@@ -1,0 +1,37 @@
+#!/bin/bash
+
+bsub -W 180 -n 1 Rscript results_abiotic.r
+for species in 'Amerikaanse eik' 'Asperge_Liggende asperge' 'Bevertjes' 'Bochtige smele' 'Bosaardbei' 'Bosanemoon' 'Brede stekelvaren' 'Brede wespenorchis_Duinwespenorchis' 'Dalkruid' 'Duizendblad' 'Eenarig wollegras' 'Engels raaigras' 'Es' 'Gestreepte witbol' 'Gevlekte orchis_Bosorchis' 'Gewone dophei' 'Gewone vogelkers' 'Gewoon kweldergras' 'Gewoon reukgras' 'Grote brandnetel' 'Hazelaar' 'Hulst' 'Jakobskruiskruid_Duinkruiskruid' 'Kleine ratelaar' 'Kleine veenbes' 'Kleine zonnedauw' 'Knolboterbloem' 'Kraaihei' 'Kruipwilg' 'Moeraswespenorchis' 'Muurpeper' 'Parnassia' 'Peen' 'Pitrus' 'Ratelpopulier' 'Ruwe berk' 'Ruwe smele' 'Slanke sleutelbloem' 'Smalle stekelvaren' 'Spaanse aak' 'Tengere rus' 'Tweerijige zegge' 'Veenpluis' 'Watermunt' 'Wilde gagel' 'Witte klaverzuring' 'Zachte berk' 'Zandstruisgras' 'Zoete kers' 'Zomereik'
+do
+  bsub -W 1440 -n 1 Rscript results_certain.r "$species"
+  bsub -W 1440 -n 1 Rscript results_uncertain.r "$species"
+  sleep 1
+done
+
[…]

### 2024-03-18 paper_codes.Rmd
update

@@
 title: "Paper Codes"
-author: "Markus Viljanen"
+author: "Makrus Viljanen"
 date: "2023-08-09"
@@
 
+##Plot grid
+
+```{r}
+colors = c("#1f77b4","#64a9d9","#bfe4ff",
+          "#ff7f0e","#ffae66","#ffddbf",
+          "#2ca02c","#6acf6a","#bfffbf",
+          "#d62627","#ea6d6e","#ffbfbf",
+          "#9466bd","#b88fde","#e0bfff")
+names(colors) = c('Closed estuaries',
+ 'Dunes mainland',
+ 'Dunes Wadden islands',
+ 'Higher sandy soils East',
+ 'Higher sandy soils North',
+ 'Higher sandy soils South',
+ 'Higher sandy soils Veluwe',
+ 'Hill country Limburg ',
+ 'Low peat Holland',
+ 'Low peat North',
+ 'River area ',
+ 'Sea clay middle',
+ 'Sea clay North',
+ 'Sea clay South ',
+ 'Tidal areas Zeeland')
+ggplot(NL) + geom_sf() +
+	geom_sf(aes(geometry=geometry.1, fill=lmf_english,linewidth=0.1,colour=""), data=sf.grid) +
+  scale_linewidth_identity() +
+  scale_fill_manual(values=colors, na.value='gray',name='Physical Geological Region') +
+  scale_colour_manual(values=NA) +
+  guides(colour=guide_legend("Not mappable", override.aes=list(colour="gray"))) +
+  scale_y_continuous(breaks = c(-100,0, 100), labels = c("363000","463000", "563000")) +
+  scale_x_continuous(breaks = c(-100,0, 100), labels = c("55000","155000", "255000")) +
+	theme(legend.position="right") #+ggtitle('Grid of Netherlands: Landuse') 
+ggsave('results_paper/fgr.png', width=1600, heigh=1500, units='px')
+```
+
+```{r}
+prov_geom <- read_sf("provincies.gpkg")$geom
+prov_geom <- (prov_geom - matrix(data=c(155000, 463000), ncol=2)) * matrix(data=c(1/1000, 1/1000), ncol=2)
+colors = c("#1f77b4","#aec7e8",
+          "#ff7f0e","#ffba78",
+          "#2ca02c","#98df8a",
+          "#d62627","#ff9896",
+          "#9466bd","#c5b0d5",
+          "#8c564b","#c49c94")
+names(colors) = c('Drenthe',
+ 'Flevoland',
+ 'Fryslân',
+ 'Gelderland',
+ 'Groningen',
+ 'Limburg',
+ 'Noord-Brabant',
+ 'Noord-Holland',
+ 'Overijssel',
+ 'Utrecht',
+ 'Zeeland',
+ 'Zuid-Holland')
+ggplot() + 
+	geom_sf(aes(geometry=geometry.1, fill=naam,linewidth=0.1), data=sf.grid) +
+  geom_sf(aes(geometry=prov_geom),fill=NA) +
+  scale_linewidth_identity() +
+  scale_fill_manual(values=colors, na.value='gray',name='Province') +
+  scale_y_continuous(breaks = c(-100,0, 100), labels = c("363000","463000", "563000")) +
+  scale_x_continuous(breaks = c(-100,0, 100), labels = c("55000","155000", "255000")) +
+	theme(legend.position="right") #+ggtitle('Grid of Netherlands: Landuse') 
+ggsave('results_paper/provinces.png', width=1600, heigh=1200, units='px')
+
+```
 ## Plot field visits
@@
 # Linear terms
+labels <- c("Organic Material","Carbon/Nitrogen","Nitrogen","Phosphor","Potassium")
 terms.plot <- terms %>% filter((term == "fixed") & (ID != "intercept")) %>% 
@@
   scale_color_discrete(name = "Model Abiotics") +
-  scale_x_discrete(name='Covariate') + 
+  scale_x_discrete(name='Covariate',labels= labels) + 
   scale_y_continuous(name='Value') + #coord_cartesian(ylim=c(-5,5))+ #ylim=c(-10,10)
@@
   mutate(term=recode(term, "spatial" = "Spatial (km)"))%>%
-  mutate(meta=factor(recode(meta, "Certain" = "Two-Stage", "Uncertain"="Joint"), levels=c("Two-Stage", "Joint")))
+  mutate(meta=factor(recode(meta, "Certain" = "Two-Stage", "Uncertain"="Joint"), levels=c("Two-Stage", "Joint"))) %>%
+  mutate(model=recode(model, "ORG_STOF.z"="Organic Material","CN.z"="Carbon/Nitrogen","N_totaal.z"="Nitrogen","P_totaal.z"="Phosphor","K_CaCl2.z"="Potassium"))
 plot.extrapolate <- ggplot(terms.plot, aes(x=ID, y=mean, col=meta, group=meta)) + 
@@
 ```{r}
-# Leave province out
-sf.grid$provincie <- as.factor(sf.grid$naam)
-plot.validation.1 <- ggplot(sf.grid) +
-	geom_sf(aes(geometry=geometry.1, fill=provincie)) + 
-	geom_sf(fill=NA, data=NL) +
-  scale_fill_discrete(name = "Province") +
-	ggtitle('Leave-province-out')
-
-# Leave FGR region out
-sf.grid$fgr <- factor(sf.grid$lmf_indeling, levels=c("afz", "duo", "duw", "hll", "hzn", "hzo", "hzv",
-                                                     "hzz", "lvh", "lvn", "riv", "zkm", "zkn", "zkz"))
-plot.validation.2 <- ggplot(sf.grid) +
-	geom_sf(aes(geometry=geometry.1, fill=fgr)) + 
-	geom_sf(fill=NA, data=NL) +
-  scale_fill_manual(name = "FGR", values=c("afz"="lightseagreen", "duo"="darkgoldenrod", "duw"="darkgoldenrod1", "hll"="forestgreen", 
-                                           "hzn"="wheat1", "hzo"="wheat2", "hzv"="wheat3", "hzz"="wheat4", 
-                                           "lvh"="darkseagreen", "lvn"="darkseagreen1", "riv"="steelblue", 
-                                           "zkm"="snow2", "zkn"="snow3", "zkz"="snow4")) +
-	ggtitle('Leave-FGR-out')
+colors = c("#1f77b4","#64a9d9","#bfe4ff",
+          "#ff7f0e","#ffae66","#ffddbf",
+          "#2ca02c","#6acf6a","#bfffbf",
+          "#d62627","#ea6d6e","#ffbfbf",
+          "#9466bd","#b88fde","#e0bfff")
+names(colors) = c('Closed estuaries',
+ 'Dunes mainland',
+ 'Dunes Wadden islands',
+ 'Higher sandy soils East',
+ 'Higher sandy soils North',
+ 'Higher sandy soils South',
+ 'Higher sandy soils Veluwe',
+ 'Hill country Limburg ',
+ 'Low peat Holland',
+ 'Low peat North',
+ 'River area ',
+ 'Sea clay middle',
+ 'Sea clay North',
+ 'Sea clay South ',
+ 'Tidal areas Zeeland')
+plot.validation.2 <- ggplot(NL) + geom_sf() +
+	geom_sf(aes(geometry=geometry.1, fill=lmf_english,linewidth=0.1,colour=""), data=sf.grid) +
+  scale_linewidth_identity() +
+  scale_fill_manual(values=colors, na.value='gray',name='Physical Geological Region') +
+  scale_colour_manual(values=NA) +
+  guides(colour=guide_legend("Not mappable", override.aes=list(colour="gray"))) +
+  scale_y_continuous(breaks = c(-100,0, 100), labels = c("363000","463000", "563000")) +
+  scale_x_continuous(breaks = c(-100,0, 100), labels = c("55000","155000", "255000")) +
+	theme(legend.position="right") +ggtitle('Leave-FGR-out')
+
+prov_geom <- read_sf("provincies.gpkg")$geom
+prov_geom <- (prov_geom - matrix(data=c(155000, 463000), ncol=2)) * matrix(data=c(1/1000, 1/1000), ncol=2)
+colors = c("#1f77b4","#aec7e8",
+          "#ff7f0e","#ffba78",
+          "#2ca02c","#98df8a",
+          "#d62627","#ff9896",
+          "#9466bd","#c5b0d5",
+          "#8c564b","#c49c94")
+names(colors) = c('Drenthe',
+ 'Flevoland',
+ 'Fryslân',
+ 'Gelderland',
+ 'Groningen',
+ 'Limburg',
+ 'Noord-Brabant',
+ 'Noord-Holland',
+ 'Overijssel',
+ 'Utrecht',
+ 'Zeeland',
+ 'Zuid-Holland')
+plot.validation.1 <- ggplot() + 
+	geom_sf(aes(geometry=geometry.1, fill=naam,linewidth=0.1), data=sf.grid) +
+  geom_sf(aes(geometry=prov_geom),fill=NA) +
+  scale_linewidth_identity() +
+  scale_fill_manual(values=colors, na.value='gray',name='Province') +
+  scale_y_continuous(breaks = c(-100,0, 100), labels = c("363000","463000", "563000")) +
+  scale_x_continuous(breaks = c(-100,0, 100), labels = c("55000","155000", "255000")) +
+	theme(legend.position="right") +ggtitle('Leave-province-out')
 
@@
 plot.all <- ggdraw() +
-  draw_plot(plot.validation.1, x = 0, y = 0, width = 0.53, height = 1) +
-  draw_plot(plot.validation.2, x = 0.53, y = 0, width = 0.47, height = 1) +
-  draw_plot_label(label = c("A", "B"), size = 15, x = c(0.01, 0.53), y = c(0.99,0.99)) +
+  draw_plot(plot.validation.1, x = 0, y = 0, width = 0.50, height = 1) +
+  draw_plot(plot.validation.2, x = 0.49, y = 0, width = 0.55, height = 1) +
+  draw_plot_label(label = c("A", "B"), size = 15, x = c(0.01, 0.50), y = c(0.99,0.99)) +
   theme(plot.background = element_rect(fill = "white", colour = NA))
 plot.all
-ggsave('results_paper/leave_out_validation.png', width=2500, heigh=1300, units='px')
+ggsave('results_paper/leave_out_validation.png', width=4600, heigh=1800, units='px')
 

### 2024-03-18 paper_data.Rmd
update

@@
 title: "paper_data"
-author: "Markus Viljanen"
+author: "Makrus Viljanen"
 date: "2023-08-09"
@@
 ```{r}
-conn <- dbConnect(PostgreSQL(), dbname="geodata", user = "", password="", host = "geodatabase.rivm.nl")
+conn <- dbConnect(PostgreSQL(), dbname="geodata", user = "ro", password="ro!geo", host = "geodatabase.rivm.nl")
 NL <- st_read(conn, query = "SELECT * FROM nl.adm_wijkbuurt_2022_gen;") %>%
@@
 # Create data frame
-sf.grid <- data.frame(grid.circle[index], grid.rect[index]) %>% st_as_sf(crs=st_crs(NL))
+sf.grid <- data.frame( grid.rect[index]) %>% st_as_sf(crs=st_crs(NL))
 sf.grid[c("X", "Y")] <- sf.grid %>% st_geometry %>% st_centroid() %>% st_coordinates()
@@
 sf.grid <- sf.grid %>% st_join(provincie , left=T, st_within)
-#plot(provincie)
+# plot(provincie)
 ```

### 2024-03-18 simulation.R

@@ -1,0 +1,449 @@
+library(INLA)
+library(sf)
+library(ggplot2)
+library(colorspace)
+library(cowplot)
+library(dplyr)
+
+fn <- "results_simulation_new_4.csv"
+if (file.exists(fn)) file.remove(fn)
+
+# Simulation parameters
+for (i in seq(1,100)) {
+  print(sprintf("   Simulation %d", i))
+  #i <- 1
+  
+  # Simulation parameters
+  sd.obs <- sqrt(0.3)
+  sd.spde <- sqrt(0.5)
+  N1 <- 100
+  N2 <- 1000
+  
+  # True effects
+  b.0 <- -2.0
+  b.Z1 <- 1.0
+  b.Z2 <- 1.0
+  
+  # Create mesh
+  limits <- cbind(c(0,1,1,0,0), c(0,0,1,1,0))
+  mesh <- inla.mesh.2d(loc.domain=limits, cutoff=0.03, max.edge=c(0.03,0.12), offset=-0.03) #0.02 0.07
+  
+  # Create grid
+  sf.boundary <- st_sfc(st_polygon(list(rbind(c(0,0), c(1,0), c(1,1), c(0,1), c(0,0)))))
+  sf.rectangles <- st_make_grid(sf.boundary, cellsize = .02, square = T)
+  sf.grid <- st_sf(geometry=sf.rectangles)
+  sf.grid[c("X", "Y")] <- sf.grid %>% st_geometry %>% st_centroid() %>% st_coordinates()
+  
+  # Base R plot of grid and mesh
+  plot(sf.boundary, col='blue')
+  plot(sf.rectangles, add=T)
+  plot(mesh, add=T)
+  
+  # Define random field (Z)
+  range0 <- 0.3
+  sigma0 <- sqrt(sd.spde)
+  kappa0 <- sqrt(8)/range0
+  tau0 <- 1/(sqrt(4*pi)*kappa0*sigma0)
+  
+  spde <- inla.spde2.matern(mesh,
+                            B.tau=matrix(c(log(tau0), -1, +1),nrow=1,ncol=3),
+                            B.kappa=matrix(c(log(kappa0), 0, -1),nrow=1,ncol=3), 
+                            theta.prior.mean=c(0, 0),
+                            theta.prior.prec=c(0.1, 0.1))
+  mesh.index <- inla.spde.make.index(name="field", n.spde=spde$n.spde)
+  
[…]

### 2024-03-18 simulation.R
Delete simulation.R

@@ -1,449 +1,0 @@
-library(INLA)
-library(sf)
-library(ggplot2)
-library(colorspace)
-library(cowplot)
-library(dplyr)
-
-fn <- "results_simulation_new_4.csv"
-if (file.exists(fn)) file.remove(fn)
-
-# Simulation parameters
-for (i in seq(1,100)) {
-  print(sprintf("   Simulation %d", i))
-  #i <- 1
-  
-  # Simulation parameters
-  sd.obs <- sqrt(0.3)
-  sd.spde <- sqrt(0.5)
-  N1 <- 100
-  N2 <- 1000
-  
-  # True effects
-  b.0 <- -2.0
-  b.Z1 <- 1.0
-  b.Z2 <- 1.0
-  
-  # Create mesh
-  limits <- cbind(c(0,1,1,0,0), c(0,0,1,1,0))
-  mesh <- inla.mesh.2d(loc.domain=limits, cutoff=0.03, max.edge=c(0.03,0.12), offset=-0.03) #0.02 0.07
-  
-  # Create grid
-  sf.boundary <- st_sfc(st_polygon(list(rbind(c(0,0), c(1,0), c(1,1), c(0,1), c(0,0)))))
-  sf.rectangles <- st_make_grid(sf.boundary, cellsize = .02, square = T)
-  sf.grid <- st_sf(geometry=sf.rectangles)
-  sf.grid[c("X", "Y")] <- sf.grid %>% st_geometry %>% st_centroid() %>% st_coordinates()
-  
-  # Base R plot of grid and mesh
-  plot(sf.boundary, col='blue')
-  plot(sf.rectangles, add=T)
-  plot(mesh, add=T)
-  
-  # Define random field (Z)
-  range0 <- 0.3
-  sigma0 <- sqrt(sd.spde)
-  kappa0 <- sqrt(8)/range0
-  tau0 <- 1/(sqrt(4*pi)*kappa0*sigma0)
-  
-  spde <- inla.spde2.matern(mesh,
-                            B.tau=matrix(c(log(tau0), -1, +1),nrow=1,ncol=3),
-                            B.kappa=matrix(c(log(kappa0), 0, -1),nrow=1,ncol=3), 
-                            theta.prior.mean=c(0, 0),
-                            theta.prior.prec=c(0.1, 0.1))
-  mesh.index <- inla.spde.make.index(name="field", n.spde=spde$n.spde)
-  
[…]

### 2024-04-08 paper_codes.Rmd
plots

@@
   scale_x_continuous(breaks = c(-100,0, 100), labels = c("55000","155000", "255000")) +
-	theme(legend.position="right") #+ggtitle('Grid of Netherlands: Landuse') 
+	theme(legend.position="right") +ggtitle('Physical Geographical Regions') 
 ggsave('results_paper/fgr.png', width=1600, heigh=1500, units='px')
@@
   scale_x_continuous(breaks = c(-100,0, 100), labels = c("55000","155000", "255000")) +
-	theme(legend.position="right") #+ggtitle('Grid of Netherlands: Landuse') 
+	theme(legend.position="right") +ggtitle('Provinces of the Netherlands')  
 ggsave('results_paper/provinces.png', width=1600, heigh=1200, units='px')
@@
 ```{r}
+sf.plots.rev <- sf.plots %>% arrange(has_abiotics)
 ggplot(NL) + geom_sf() +
@@
 	#geom_sf(aes(geometry=geometry), shape=3, size=0.3, data=sf.grid) + 
-	geom_sf(aes(geometry=geometry, color=has_abiotics), shape=4, size=0.5, data=sf.plots) + 
+	geom_sf(aes(geometry=geometry, color=has_abiotics), shape=4, size=0.5, data=sf.plots.rev) + 
 	theme(plot.title = element_text(hjust = 0.5), legend.title=element_text(size=10), 
 		legend.text=element_text(size=8), legend.position="right") +
-	scale_colour_manual(values=c("#377EB8", "#4DAF4A"), name="Has abiotics") #+ 
+	scale_colour_manual(values=c("#377EB8", "#4DAF4A"), name="Abiotic measurements") +
+  scale_fill_discrete(guide=guide_legend(reverse=T))
 	#ggtitle('Location of Field Visits and Abiotics') 
@@
 ```{r}
-colors = c("#1f77b4","#64a9d9","#bfe4ff",
-          "#ff7f0e","#ffae66","#ffddbf",
-          "#2ca02c","#6acf6a","#bfffbf",
-          "#d62627","#ea6d6e","#ffbfbf",
+colors = c("#1f77b4","#64a9d9",
+          "#ff7f0e","#ff9e49","#ffbd84","#ffddbf",
+          "#2ca02c",
+          "#d62627","#ea6d6e",
+          "#8c564b",
           "#9466bd","#b88fde","#e0bfff")
-names(colors) = c('Closed estuaries',
- 'Dunes mainland',
- 'Dunes Wadden islands',
- 'Higher sandy soils East',
- 'Higher sandy soils North',
- 'Higher sandy soils South',
- 'Higher sandy soils Veluwe',
+names(colors) = c(
+ 'Dunes mainland', 'Dunes Wadden islands',
+ 'Higher sandy soils East', 'Higher sandy soils North', 'Higher sandy soils South', 'Higher sandy soils Veluwe',
  'Hill country Limburg ',
- 'Low peat Holland',
- 'Low peat North',
+ 'Low peat Holland', 'Low peat North',
  'River area ',
- 'Sea clay middle',
- 'Sea clay North',
- 'Sea clay South ',
- 'Tidal areas Zeeland')
+ 'Sea clay middle', 'Sea clay North', 'Sea clay South ')
 plot.validation.2 <- ggplot(NL) + geom_sf() +
-	geom_sf(aes(geometry=geometry.1, fill=lmf_english,linewidth=0.1,colour=""), data=sf.grid) +
+	geom_sf(aes(geometry=geometry.1, fill=lmf_english,linewidth=0.1,colour=""), data=filter(sf.grid, lmf_english!='Tidal areas Zeeland', lmf_english!='Closed estuaries')) +
   scale_linewidth_identity() +
@@
   guides(colour=guide_legend("Not mappable", override.aes=list(colour="gray"))) +
-  scale_y_continuous(breaks = c(-100,0, 100), labels = c("363000","463000", "563000")) +
-  scale_x_continuous(breaks = c(-100,0, 100), labels = c("55000","155000", "255000")) +
+  # scale_y_continuous(breaks = c(-100,0, 100), labels = c("363000","463000", "563000")) +
+  # scale_x_continuous(breaks = c(-100,0, 100), labels = c("55000","155000", "255000")) +
 	theme(legend.position="right") +ggtitle('Leave-FGR-out')
@@
   scale_fill_manual(values=colors, na.value='gray',name='Province') +
-  scale_y_continuous(breaks = c(-100,0, 100), labels = c("363000","463000", "563000")) +
-  scale_x_continuous(breaks = c(-100,0, 100), labels = c("55000","155000", "255000")) +
+  # scale_y_continuous(breaks = c(-100,0, 100), labels = c("363000","463000", "563000")) +
+  # scale_x_continuous(breaks = c(-100,0, 100), labels = c("55000","155000", "255000")) +
 	theme(legend.position="right") +ggtitle('Leave-province-out')
@@
 plot.all <- ggdraw() +
-  draw_plot(plot.validation.1, x = 0, y = 0, width = 0.50, height = 1) +
-  draw_plot(plot.validation.2, x = 0.49, y = 0, width = 0.55, height = 1) +
-  draw_plot_label(label = c("A", "B"), size = 15, x = c(0.01, 0.50), y = c(0.99,0.99)) +
+  draw_plot(plot.validation.1, x = 0.01, y = 0, width = 0.47, height = 1) +
+  draw_plot(plot.validation.2, x = 0.47, y = 0, width = 0.53, height = 1) +
+  draw_plot_label(label = c("A", "B"), size = 15, x = c(0.0, 0.455), y = c(0.99,0.99)) +
   theme(plot.background = element_rect(fill = "white", colour = NA))
 plot.all
-ggsave('results_paper/leave_out_validation.png', width=4600, heigh=1800, units='px')
+ggsave('results_paper/leave_out_validation.png', width=3400, heigh=1400, units='px')
 

### 2024-04-08 simulation.R

@@ -1,15 +1,19 @@
+#author: "Markus Viljanen"
+#date: "2024-03-19"
+
 library(INLA)
 library(sf)
 library(ggplot2)
 library(colorspace)
 library(cowplot)
 library(dplyr)
+library(tidyr)
 
-fn <- "results_simulation_new_4.csv"
+fn <- "results_simulation.csv"
 if (file.exists(fn)) file.remove(fn)
 
 # Simulation parameters
-for (i in seq(1,100)) {
+for (i in seq(1,200)) {
   print(sprintf("   Simulation %d", i))
   #i <- 1
   
@@ -26,7 +30,7 @@ for (i in seq(1,100)) {
   
   # Create mesh
   limits <- cbind(c(0,1,1,0,0), c(0,0,1,1,0))
-  mesh <- inla.mesh.2d(loc.domain=limits, cutoff=0.03, max.edge=c(0.03,0.12), offset=-0.03) #0.02 0.07
+  mesh <- inla.mesh.2d(loc.domain=limits, cutoff=0.03, max.edge=c(0.03,0.12), offset=-0.03) 
   
   # Create grid
   sf.boundary <- st_sfc(st_polygon(list(rbind(c(0,0), c(1,0), c(1,1), c(0,1), c(0,0)))))
@@ -77,11 +81,6 @@ for (i in seq(1,100)) {
   xy.obs <- data.frame(X=sample(1:N / N), Y=sample(1:N / N))
   
   # Run repeated samplings of the given simulation
-  #spatial.misalignment <- F
-  #measurement.error <- F
-  #confounded <- T
-  #control.z2 <- T
-  #spatial <- T
   for (spatial.misalignment in c(F,T)) {
     for (measurement.error in c(F,T)) {
       for (confounded in c(T)) {#c(F,T)
@@ -374,7 +373,7 @@ ggsave('results_paper/simulation.png', width=3000, heigh=1200, units='px')
 # Plot illustration of Kriging: Berkson & Classical type errors
 
 # True values of covariate Z at grid points
-Y.choice <- 0.90 #0.75
+Y.choice <- 0.80 #0.75
 sf.line <- data.frame(X=seq(0,1,0.001), Y=Y.choice)
 A.line <- inla.spde.make.A(mesh, loc=as.matrix(sf.line))
 # True Z1 and observation standard error
[…]

### 2024-04-08 simulation_results.R

@@ -1,0 +1,48 @@
+#author: "Markus Viljanen"
+#date: "2024-03-19"
+library(ggplot2)
+library(dplyr)
+library(tidyr)
+
+fn <- "results_simulation.csv"
+df <- read.table(fn, header=T, sep=",", row.names=NULL)
+cov.true <- 1.0
+df$true <- cov.true
+
+# Results from first simulation
+df %>% filter((Simulation == 1) & Confounded & Control.z2) %>% 
+  select(Spatial.Misalignment, Measurement.Error, 
+         Model, mean, sd, rmse.y, rmse.z1)
+
+# Summary of simulations: coefficient and predictive accuracy
+df %>% filter(Confounded & Control.z2) %>% #%>% filter(Simulation <= 100) 
+  group_by(Spatial.Misalignment, Measurement.Error, Model) %>% 
+  summarize(n=n(), 
+            mean.coef=round(mean(mean,na.rm=T),2), 
+            mean.coef.sd=round(mean(sd,na.rm=T),2),#sd.coef=round(sd(mean),2), 
+            #coverage=round(mean((true >= X0.025quant) & (true <= X0.975quant)),2),
+            rmse.coef=round(mean(sqrt((true-mean)**2), na.rm=T),2),  
+            rmse.y=round(mean(rmse.y),3), 
+            rmse.z1=round(mean(rmse.z1),3)) %>%
+  write.csv(file="results_paper/results_simulation.csv", row.names=F)
+
+# Summary of simulations: estimated coefficient of X1 in two models
+temp <- df %>% filter(Confounded & Control.z2 & Spatial.Misalignment & Measurement.Error) %>%
+  select(Simulation, Model, mean) %>% pivot_wider(names_from=Model, values_from=mean)
+temp %>% ggplot(aes(x=`Two-Stage`, y=Joint)) + geom_point() + geom_abline() +
+  geom_smooth(method='lm', formula= y~x)
+
+# Plot estimated coefficients from joint vs. two-stage model
+df.subset <- df %>% filter(Confounded & Control.z2 & Spatial.Misalignment & Measurement.Error) %>% 
+  mutate(Model = factor(Model, levels=c("Two-Stage", "Joint")))
+plot <- ggplot(df.subset, aes(x=Simulation, y=mean, col=Model, group=Model)) + 
+  geom_point(position=position_dodge(width=0.5), size=2) +
[…]

### 2024-04-26 paper_codes.Rmd
final changes

--- 2023-08-23 paper_codes.Rmd
@@
 These coordinates are calculated from the center of the Netherlands in meters.
-For simplicity, we convert them to kilometers below.
+We convert them to kilometers from the center below.
 
@@
 Data for vegetation plots and abiotic measurements data
-Coordinates are in EPSG:28992, the Projected coordinate system for Netherlands.
 
@@
 Data for Netherlands grid
-Coordinates are in EPSG:28992, the Projected coordinate system for Netherlands.
 
@@
 ggsave('results_paper/landuse.png', width=1600, heigh=1200, units='px')
+ggsave('results_paper/landuse.tiff', width=1600, heigh=1200, units='px', compression="lzw")
+
 ```
@@
 ggsave('results_paper/landuse_vs_visit.png', width=1400, heigh=900, units='px')
+ggsave('results_paper/landuse_vs_visit.tiff', width=1400, heigh=900, units='px', compression="lzw")
 ```
@@
 	#geom_sf(aes(geometry=geometry), shape=3, size=0.3, data=sf.grid) + 
-	geom_sf(aes(geometry=geometry, color=has_abiotics), shape=4, size=0.5, data=sf.plots) + 
+	geom_sf(aes(geometry=geometry, color=has_abiotics), shape=4, size=0.5, 
+	        data=sf.plots %>% arrange(has_abiotics)) + 
 	theme(plot.title = element_text(hjust = 0.5), legend.title=element_text(size=10), 
 		legend.text=element_text(size=8), legend.position="right") +
-	scale_colour_manual(values=c("#377EB8", "#4DAF4A"), name="Has abiotics") #+ 
+	scale_colour_manual(values=c("#377EB8", "#4DAF4A"), name="Abiotic measurements") +
+  scale_fill_discrete(guide=guide_legend(reverse=T))
 	#ggtitle('Location of Field Visits and Abiotics') 
 ggsave('results_paper/visits.png', width=1400, heigh=1200, units='px')
-
+ggsave('results_paper/visits.tiff', width=1400, heigh=1200, units='px', compression="lzw")
 ```
@@
 ggsave('results_paper/Struikhei.png', width=2000, heigh=1400, units='px')
+ggsave('results_paper/Struikhei.tiff', width=2000, heigh=1400, units='px', compression="lzw")
 ```
@@
 ggsave('results_paper/pH.png', width=2000, heigh=1400, units='px')
+ggsave('results_paper/pH.tiff', width=2000, heigh=1400, units='px', compression="lzw")
 ```
@@
 ggsave('results_paper/pH2.png', width=3000, heigh=1400, units='px')
+ggsave('results_paper/pH2.tiff', width=3000, heigh=1400, units='px', compression="lzw")
 ```
@@
 ## Experiments: training and validation
-It is also possible to download the results from https://zenodo.org/record/8261504
-In that case, place the downloaded files in the subfolder `data_paper/` and go to the `Results: import file` section below to import the results and create the plots.
 
@@
 Combine and export the results of fitting all models to data:
-It is also possible to skip this cell and download the results from https://zenodo.org/record/8261504
-Place the data in the subfolder `data_paper/` and run the code below this cell to create the plots.
 
@@
 
-## Results: import file
+## Results
+
+It is also possible to download the results from https://zenodo.org/record/8261504
+Place the data in the subfolder `data_paper/` and run the code below this cell.
 
@@
 
-## Results: crowberry example
+## Crowberry example
 
@@
 ggsave('results_paper/species_parameters.png', width=2000, heigh=2500, units='px')
+ggsave('results_paper/species_parameters.tiff', width=2000, heigh=2500, units='px', compression="lzw")
 ```
@@
 ggsave('results_paper/species_SDM_y.png', width=3200, heigh=3000, units='px')
+ggsave('results_paper/species_SDM_y.tiff', width=3200, heigh=3000, units='px', compression="lzw")
 ```
@@
 ggsave('results_paper/species_SDM_logodds.png', width=3200, heigh=3000, units='px')
-
+ggsave('results_paper/species_SDM_logodds.tiff', width=3200, heigh=3000, units='px', compression="lzw")
 ```
@@
 ggsave(sprintf('results_paper/species_%s.png', var), width=3200, heigh=3000, units='px')
+ggsave(sprintf('results_paper/species_%s.tiff', var), width=3200, heigh=3000, units='px', compression="lzw")
 }
@@
 ggsave('results_paper/species_scatter.png', width=1800, heigh=2400, units='px')
+ggsave('results_paper/species_scatter.tiff', width=1800, heigh=2400, units='px', compression="lzw")
 ```
@@
 ## Results: all species
+
+```{r}
+results <- read.csv("data_paper/results.csv")
+```
 
@@
 ggsave('results_paper/all_abiotics.png', width=2400, heigh=4000, units='px')
+ggsave('results_paper/all_abiotics.tiff', width=2400, heigh=4000, units='px', compression="lzw")
 ```
@@
 ggsave('results_paper/species_correlation.png', width=2000, heigh=1200, units='px')
-
+ggsave('results_paper/species_correlation.tiff', width=2000, heigh=1200, units='px', compression="lzw")
 ```
@@
 Validation set illustration
+
+```{r}
+
+sf.grid$fgr <- factor(sf.grid$lmf_indeling, levels=c("afz", "duo", "duw", "hll", "hzn", "hzo", "hzv",
+                                                     "hzz", "lvh", "lvn", "riv", "zkm", "zkn", "zkz"))
+levels(sf.grid$fgr) <- list("Closed estuaries"="afz",
+                         "Dunes mainland"="duo",
+                         "Dunes Wadden islands"="duw",
+                         "Tidal areas Wadden islands"="gtw",
+                         "Tidal areas Zeeland"="gtz",
+                         "Hill country Limburg"="hll",
+                         "Higher sandy soils North"="hzn",
+                         "Higher sandy soils East"="hzo",
+                         "Higher sandy soils Veluwe"="hzv",
+                         "Higher sandy soils South"="hzz",
+                         "Low peat Holland"="lvh",
+                         "Low peat North"="lvn",
+                         "River area"="riv",
+                         "Sea clay middle"="zkm",
+                         "Sea clay North"="zkn",
+                         "Sea clay South"="zkz")
+
+colors = c("Closed estuaries"="#d62627",
+           "Dunes mainland"="#ea6d6e",
+           "Dunes Wadden islands"="#ffbfbf",
+           "Tidal areas Wadden islands"="#1f77b4",
+           "Tidal areas Zeeland"="#64a9d9",
+           "Hill country Limburg"="#bfe4ff",
+           "Higher sandy soils North"="#ff7f0e",
+           "Higher sandy soils East"="#ffae66",
+           "Higher sandy soils Veluwe"="#bfffbf",
+           "Higher sandy soils South"="#ffddbf",
+           "Low peat Holland"="#6acf6a",
+           "Low peat North"="#2ca02c",
+           "River area"="#00BFC4",
+           "Sea clay middle"="#b88fde",
+           "Sea clay North"="#9466bd",
+           "Sea clay South"="#e0bfff")
+
+ggplot(NL) + geom_sf() +
+	geom_sf(aes(geometry=geometry.1, fill=fgr,linewidth=0.1,colour=""), data=sf.grid) +
+  scale_linewidth_identity() +
+  scale_fill_manual(values=colors, na.value='gray',na.translate = F,name='Physical Geological Region') +
+  scale_colour_manual(values=NA) +
+  guides(colour=guide_legend("Not mappable", override.aes=list(colour="gray"))) +
+	theme(legend.position="right") + ggtitle('Physical Geographical Regions') 
+ggsave('results_paper/fgr.png', width=2000, heigh=1400, units='px')
+ggsave('results_paper/fgr.tiff', width=2000, heigh=1400, units='px', compression="lzw")
+
+```
+
+```{r}
+prov_geom <- read_sf("data_paper/provincies.gpkg")$geom
+prov_geom <- (prov_geom - matrix(data=c(155000, 463000), ncol=2)) * matrix(data=c(1/1000, 1/1000), ncol=2)
+colors = c("#1f77b4","#aec7e8",
+          "#ff7f0e","#ffba78",
+          "#2ca02c","#98df8a",
+          "#d62627","#ff9896",
+          "#9466bd","#c5b0d5",
+          "#8c564b","#c49c94")
+names(colors) = c('Drenthe',
+ 'Flevoland',
+ 'Fryslân',
+ 'Gelderland',
+ 'Groningen',
+ 'Limburg',
+ 'Noord-Brabant',
+ 'Noord-Holland',
+ 'Overijssel',
+ 'Utrecht',
+ 'Zeeland',
+ 'Zuid-Holland')
+ggplot() + 
+	geom_sf(aes(geometry=geometry.1, fill=naam,linewidth=0.1), data=sf.grid) +
+  geom_sf(aes(geometry=prov_geom),fill=NA) +
+  scale_linewidth_identity() +
+  scale_fill_manual(values=colors, na.value='gray',name='Province') +
+  theme(legend.position="right") +ggtitle('Provinces of the Netherlands')  
+ggsave('results_paper/provinces.png', width=1700, heigh=1200, units='px')
+ggsave('results_paper/provinces.tiff', width=1700, heigh=1200, units='px', compression="lzw")
+
+```
 
@@
 ggsave('results_paper/leave_out_validation.png', width=2500, heigh=1300, units='px')
-
+ggsave('results_paper/leave_out_validation.tiff', width=2500, heigh=1300, units='px', compression="lzw")
 ```
@@
 ggsave('results_paper/rmse_comparison.png', width=1200, heigh=800, units='px')
-
+ggsave('results_paper/rmse_comparison.tiff', width=1200, heigh=800, units='px', compression="lzw")
 ```
@@
 ggsave('results_paper/all_prevalences.png', width=2800, heigh=4000, units='px')
+ggsave('results_paper/all_prevalences.tiff', width=2800, heigh=4000, units='px', compression="lzw")
 ```
@@
 ggsave('results_paper/all_prevalences2.png', width=2800, heigh=4000, units='px')
-```
+ggsave('results_paper/all_prevalences2.tiff', width=2800, heigh=4000, units='px', compression="lzw")
+```

### 2024-04-26 simulation.R

@@ -13,9 +13,8 @@ fn <- "results_simulation.csv"
 if (file.exists(fn)) file.remove(fn)
 
 # Simulation parameters
-for (i in seq(1,200)) {
+for (i in seq(1,200)) {  #i <- 1
   print(sprintf("   Simulation %d", i))
-  #i <- 1
   
   # Simulation parameters
   sd.obs <- sqrt(0.3)
@@ -368,6 +367,7 @@ plot.all <- ggdraw() +
   theme(plot.background = element_rect(fill = "white", colour = NA))
 plot.all
 ggsave('results_paper/simulation.png', width=3000, heigh=1200, units='px')
+ggsave('results_paper/simulation.tiff', width=3000, heigh=1200, units='px', compression="lzw")
 
 
 # Plot illustration of Kriging: Berkson & Classical type errors
@@ -428,6 +428,7 @@ plot.line <- ggplot(sf.slice) + geom_line(aes(x=X,y=Value, col=Covariate)) +
   xlab(expression(r))
 plot.line
 ggsave('results_paper/simulation_kriging.png', width=1500, heigh=900, units='px')
+ggsave('results_paper/simulation_kriging.tiff', width=1500, heigh=900, units='px', compression="lzw")
 
 
 # Load saved results
@@ -441,7 +442,7 @@ df %>% filter((Simulation == 1) & Confounded & Control.z2) %>%
          Model, mean, sd, rmse.y, rmse.z1)
 
 # Summary of simulations: coefficient and predictive accuracy
-df %>% filter(Simulation <= 100) %>% filter(Confounded & Control.z2) %>%
+df %>% filter(Confounded & Control.z2) %>%
   group_by(Spatial.Misalignment, Measurement.Error, Model) %>% 
   summarize(n=n(), 
             mean.coef=round(mean(mean,na.rm=T),2), 
@@ -453,7 +454,7 @@ df %>% filter(Simulation <= 100) %>% filter(Confounded & Control.z2) %>%
 
 # Plot estimated coefficients from joint vs. two-stage model
 df.subset <- df %>% filter(Confounded & Control.z2 & Spatial.Misalignment & Measurement.Error) %>% 
-  mutate(Model = factor(Model, levels=c("Two-Stage", "Joint")))
+  mutate(Model = factor(Model, levels=c("Two-Stage", "Joint"))) %>% filter(Simulation <= 100)
[…]

### 2024-04-26 simulation_results.R

@@ -1,48 +1,0 @@
-#author: "Markus Viljanen"
-#date: "2024-03-19"
-library(ggplot2)
-library(dplyr)
-library(tidyr)
-
-fn <- "results_simulation.csv"
-df <- read.table(fn, header=T, sep=",", row.names=NULL)
-cov.true <- 1.0
-df$true <- cov.true
-
-# Results from first simulation
-df %>% filter((Simulation == 1) & Confounded & Control.z2) %>% 
-  select(Spatial.Misalignment, Measurement.Error, 
-         Model, mean, sd, rmse.y, rmse.z1)
-
-# Summary of simulations: coefficient and predictive accuracy
-df %>% filter(Confounded & Control.z2) %>% #%>% filter(Simulation <= 100) 
-  group_by(Spatial.Misalignment, Measurement.Error, Model) %>% 
-  summarize(n=n(), 
-            mean.coef=round(mean(mean,na.rm=T),2), 
-            mean.coef.sd=round(mean(sd,na.rm=T),2),#sd.coef=round(sd(mean),2), 
-            #coverage=round(mean((true >= X0.025quant) & (true <= X0.975quant)),2),
-            rmse.coef=round(mean(sqrt((true-mean)**2), na.rm=T),2),  
-            rmse.y=round(mean(rmse.y),3), 
-            rmse.z1=round(mean(rmse.z1),3)) %>%
-  write.csv(file="results_paper/results_simulation.csv", row.names=F)
-
-# Summary of simulations: estimated coefficient of X1 in two models
-temp <- df %>% filter(Confounded & Control.z2 & Spatial.Misalignment & Measurement.Error) %>%
-  select(Simulation, Model, mean) %>% pivot_wider(names_from=Model, values_from=mean)
-temp %>% ggplot(aes(x=`Two-Stage`, y=Joint)) + geom_point() + geom_abline() +
-  geom_smooth(method='lm', formula= y~x)
-
-# Plot estimated coefficients from joint vs. two-stage model
-df.subset <- df %>% filter(Confounded & Control.z2 & Spatial.Misalignment & Measurement.Error) %>% 
-  mutate(Model = factor(Model, levels=c("Two-Stage", "Joint")))
-plot <- ggplot(df.subset, aes(x=Simulation, y=mean, col=Model, group=Model)) + 
-  geom_point(position=position_dodge(width=0.5), size=2) +
[…]
